---
title: "Advanced SERO: Parameter Optimization and Custom Analysis"
author: "Prince Oppong Boakye"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced SERO: Parameter Optimization and Custom Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

This vignette demonstrates advanced usage of the SERO package, including parameter optimization, sensitivity analysis, and custom analysis workflows for different scenarios.

```{r setup}
library(SERO)
```

## Parameter Sensitivity Analysis

### Buffer Size Sensitivity for Hotspot Detection

```{r}
# Load data
data <- sero_load_data()

# Test different buffer sizes
buffer_values <- c(300, 500, 800, 1000, 1200, 1500)
buffer_results <- data.frame(
  buffer = buffer_values,
  hotspots = integer(length(buffer_values)),
  max_density = numeric(length(buffer_values))
)

for (i in seq_along(buffer_values)) {
  result <- sero_identify_hotspots(data$accident, buffer = buffer_values[i])
  buffer_results$hotspots[i] <- nrow(result$hotspots)
  if (nrow(result$hotspots) > 0) {
    buffer_results$max_density[i] <- max(result$hotspots$density, na.rm = TRUE)
  }
}

print(buffer_results)
```

### Weight Sensitivity for Optimal Locations

```{r}
# Test different weight combinations
weight_scenarios <- list(
  "Accident-focused" = list(
    weight_accidents = 0.6, weight_population = 0.15, weight_roads = 0.1,
    weight_landuse = 0.05, weight_distance = 0.05, weight_distribution = 0.05
  ),
  "Population-focused" = list(
    weight_accidents = 0.2, weight_population = 0.5, weight_roads = 0.15,
    weight_landuse = 0.05, weight_distance = 0.05, weight_distribution = 0.05
  ),
  "Balanced" = list(
    weight_accidents = 0.3, weight_population = 0.25, weight_roads = 0.2,
    weight_landuse = 0.1, weight_distance = 0.1, weight_distribution = 0.05
  )
)

weight_results <- data.frame(
  scenario = names(weight_scenarios),
  locations = integer(length(weight_scenarios)),
  avg_score = numeric(length(weight_scenarios))
)

for (i in seq_along(weight_scenarios)) {
  params <- weight_scenarios[[i]]
  result <- do.call(sero_calculate_optimal_locations, c(list(data = data, max_locations = 5), params))
  weight_results$locations[i] <- nrow(result$locations)
  if (nrow(result$locations) > 0) {
    weight_results$avg_score[i] <- mean(result$locations$total_score, na.rm = TRUE)
  }
}

print(weight_results)
```

## Scenario-Based Analysis

### Urban Dense Area Analysis

```{r}
# Parameters optimized for dense urban areas
urban_analysis <- sero_analyze(
  risk_categories = c(1, 2),
  max_locations = 8,        # More locations for dense coverage
  max_routes = 15
)

# Custom hotspot analysis for urban areas
urban_hotspots <- sero_identify_hotspots(
  data$accident,
  buffer = 600,             # Smaller buffer for precise urban hotspots
  min_events = 3            # Lower threshold due to higher density
)

cat("Urban Analysis Results:\n")
cat("- Hotspots identified:", nrow(urban_hotspots$hotspots), "\n")
cat("- Optimal locations:", nrow(urban_analysis$locations$locations), "\n")
cat("- Routes calculated:", nrow(urban_analysis$routes$routes), "\n")
```

### Rural/Suburban Area Analysis

```{r}
# Parameters optimized for rural/suburban areas
suburban_hotspots <- sero_identify_hotspots(
  data$accident,
  buffer = 1500,            # Larger buffer for broader coverage
  min_events = 6            # Higher threshold for significance
)

suburban_locations <- sero_calculate_optimal_locations(
  data,
  max_locations = 4,        # Fewer locations for broader coverage
  weight_accidents = 0.4,
  weight_population = 0.15,
  weight_roads = 0.25,      # Higher weight on road accessibility
  weight_landuse = 0.1,
  weight_distance = 0.05,
  weight_distribution = 0.05
)

cat("Suburban Analysis Results:\n")
cat("- Hotspots identified:", nrow(suburban_hotspots$hotspots), "\n")
cat("- Optimal locations:", nrow(suburban_locations$locations), "\n")
```

## Multi-Criteria Decision Analysis

### Understanding Criteria Weights

```{r}
# Demonstrate the impact of different criteria
criteria_tests <- list(
  "Equal weights" = list(
    weight_accidents = 1/6, weight_population = 1/6, weight_roads = 1/6,
    weight_landuse = 1/6, weight_distance = 1/6, weight_distribution = 1/6
  ),
  "Safety-focused" = list(
    weight_accidents = 0.5, weight_population = 0.1, weight_roads = 0.15,
    weight_landuse = 0.1, weight_distance = 0.1, weight_distribution = 0.05
  ),
  "Coverage-focused" = list(
    weight_accidents = 0.2, weight_population = 0.4, weight_roads = 0.1,
    weight_landuse = 0.05, weight_distance = 0.05, weight_distribution = 0.2
  )
)

for (scenario_name in names(criteria_tests)) {
  cat("\n", scenario_name, ":\n")
  params <- criteria_tests[[scenario_name]]
  result <- do.call(sero_calculate_optimal_locations, c(list(data = data, max_locations = 3), params))
  
  if (nrow(result$locations) > 0) {
    cat("- Locations found:", nrow(result$locations), "\n")
    cat("- Average total score:", round(mean(result$locations$total_score), 3), "\n")
    cat("- Score range:", round(min(result$locations$total_score), 3), "-", 
        round(max(result$locations$total_score), 3), "\n")
  }
}
```

## Custom Analysis Workflows

### Emergency Response Time Analysis

```{r}
# Calculate comprehensive response analysis
optimal_locs <- sero_calculate_optimal_locations(data, max_locations = 5)
all_routes <- sero_calculate_routes(optimal_locs, data$accident, max_routes = 50)

# Analyze response times
response_stats <- list(
  mean_time = mean(all_routes$routes$estimated_time),
  median_time = median(all_routes$routes$estimated_time),
  max_time = max(all_routes$routes$estimated_time),
  under_5min = sum(all_routes$routes$estimated_time <= 5),
  under_10min = sum(all_routes$routes$estimated_time <= 10)
)

cat("Response Time Analysis:\n")
cat("- Mean response time:", round(response_stats$mean_time, 1), "minutes\n")
cat("- Median response time:", round(response_stats$median_time, 1), "minutes\n")
cat("- Maximum response time:", round(response_stats$max_time, 1), "minutes\n")
cat("- Accidents reachable in ≤5 min:", response_stats$under_5min, "\n")
cat("- Accidents reachable in ≤10 min:", response_stats$under_10min, "\n")
```

### Coverage Analysis by Accident Severity

```{r}
# Analyze coverage for different accident severities
if ("UKATEGORIE" %in% names(data$accident)) {
  severity_analysis <- data.frame(
    severity = c("Fatal", "Serious", "Light"),
    category = c(1, 2, 3),
    count = c(
      sum(data$accident$UKATEGORIE == 1, na.rm = TRUE),
      sum(data$accident$UKATEGORIE == 2, na.rm = TRUE),
      sum(data$accident$UKATEGORIE == 3, na.rm = TRUE)
    )
  )
  
  print(severity_analysis)
  
  # Calculate average response time by severity
  if (nrow(all_routes$routes) > 0) {
    for (sev in 1:3) {
      sev_accidents <- which(data$accident$UKATEGORIE == sev)
      sev_routes <- all_routes$routes[all_routes$routes$accident_id %in% sev_accidents[1:min(length(sev_accidents), nrow(all_routes$routes))], ]
      
      if (nrow(sev_routes) > 0) {
        cat("Severity", sev, "average response time:", 
            round(mean(sev_routes$estimated_time), 1), "minutes\n")
      }
    }
  }
}
```

## Performance Optimization

### Grid Size vs. Computation Time

```{r}
# Test different grid sizes for optimal location calculation
grid_sizes <- c(100, 150, 200, 250)
grid_performance <- data.frame(
  grid_size = grid_sizes,
  locations = integer(length(grid_sizes)),
  computation_time = numeric(length(grid_sizes))
)

for (i in seq_along(grid_sizes)) {
  start_time <- Sys.time()
  
  # Run with different grid sizes (if this parameter exists)
  result <- sero_calculate_optimal_locations(
    data,
    max_locations = 3
  )
  
  end_time <- Sys.time()
  
  grid_performance$locations[i] <- nrow(result$locations)
  grid_performance$computation_time[i] <- as.numeric(difftime(end_time, start_time, units = "secs"))
}

print(grid_performance)
```

## Quality Assessment

### Validation Metrics

```{r}
# Calculate validation metrics for optimal locations
hotspots <- sero_identify_hotspots(data$accident)
optimal_locs <- sero_calculate_optimal_locations(data, max_locations = 5)

if (nrow(hotspots$hotspots) > 0 && nrow(optimal_locs$locations) > 0) {
  # Calculate distances from optimal locations to hotspots
  hotspot_distances <- sf::st_distance(
    sf::st_transform(optimal_locs$locations, 32632),
    sf::st_transform(hotspots$hotspots, 32632)
  )
  
  # Minimum distance from each hotspot to nearest service location
  min_distances_to_hotspots <- apply(hotspot_distances, 2, min)
  
  cat("Quality Assessment:\n")
  cat("- Hotspots within 2km of service:", sum(min_distances_to_hotspots <= 2000), "/", length(min_distances_to_hotspots), "\n")
  cat("- Average distance to nearest service:", round(mean(min_distances_to_hotspots)), "meters\n")
  cat("- Maximum distance to service:", round(max(min_distances_to_hotspots)), "meters\n")
}
```

## Best Practices and Recommendations

### Parameter Selection Guidelines

1. **Buffer Size for Hotspots**:
   - Urban areas: 500-800m
   - Suburban areas: 800-1200m
   - Rural areas: 1200-2000m

2. **Multi-Criteria Weights**:
   - Safety-focused: High accident weight (0.4-0.6)
   - Coverage-focused: High population weight (0.3-0.5)
   - Balanced approach: More even distribution

3. **Number of Locations**:
   - Consider area size, population density, and budget constraints
   - Start with 3-5 locations and adjust based on coverage analysis

### Validation Steps

```{r}
# Validation workflow
cat("Recommended Validation Steps:\n")
cat("1. Check data quality and completeness\n")
cat("2. Validate hotspot identification with domain experts\n")
cat("3. Test different parameter combinations\n")
cat("4. Assess coverage of high-risk areas\n")
cat("5. Evaluate response time distributions\n")
cat("6. Consider real-world constraints (budget, infrastructure)\n")
```

## Conclusion

This advanced vignette demonstrates the flexibility and power of the SERO package for customized spatial emergency response optimization. Key takeaways:

- Parameter sensitivity analysis helps optimize results for specific scenarios
- Different weight combinations produce solutions tailored to specific objectives
- Validation metrics ensure solution quality and practical applicability
- Performance considerations help balance accuracy with computational efficiency

For additional examples and use cases, explore the package's example scripts in `inst/examples/`.
