---
title: "Introduction to SERO: Spatial Emergency Response Optimization"
author: "Prince Oppong Boakye"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to SERO: Spatial Emergency Response Optimization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

SERO (Spatial Emergency Response Optimization) is an R package designed for spatial analysis and optimization of emergency response services. This package provides comprehensive tools for:

- **Hotspot Analysis**: Identifying accident hotspots using point pattern analysis
- **Optimal Location Calculation**: Multi-criteria analysis for emergency service placement
- **Route Planning**: Calculating fastest routes to accident scenes
- **Interactive Mapping**: Real-time accident selection and route calculation

## Installation

```{r eval=FALSE}
# Install from GitHub
devtools::install_github("iprincegh/SERO")
```

```{r setup}
library(SERO)
```

## Built-in Data

SERO includes comprehensive spatial datasets for Münster, Germany:

```{r}
# Load built-in data
data <- sero_load_data()

# Explore the data structure
names(data)
```

```{r}
# Data summary
cat("Accidents:", nrow(data$accident), "\n")
cat("Population grid cells:", nrow(data$population), "\n")
cat("Road segments:", nrow(data$roads), "\n")
cat("Land use polygons:", nrow(data$landuse), "\n")
cat("Districts:", nrow(data$districts), "\n")
```

## Basic Workflow

### 1. Identify Accident Hotspots

```{r}
# Identify hotspots using point pattern analysis
hotspots <- sero_identify_hotspots(
  data$accident,
  risk_categories = c(1, 2),  # Fatal and serious accidents
  buffer = 1000,              # 1km buffer for density estimation
  min_events = 5              # Minimum 5 accidents per hotspot
)

print(hotspots)
```

```{r, fig.cap="Accident hotspots visualization"}
plot(hotspots)
```

### 2. Calculate Optimal Service Locations

The package uses 6 multi-criteria analysis criteria:

1. **Accident density proximity**
2. **Population density coverage**
3. **Road network accessibility**
4. **Land use compatibility**
5. **Distance from existing services**
6. **Spatial distribution balance**

```{r}
# Calculate optimal locations
optimal_locations <- sero_calculate_optimal_locations(
  data,
  max_locations = 3,
  weight_accidents = 0.4,
  weight_population = 0.2,
  weight_roads = 0.15,
  weight_landuse = 0.1,
  weight_distance = 0.1,
  weight_distribution = 0.05
)

print(optimal_locations)
```

```{r, fig.cap="Optimal emergency service locations"}
plot(optimal_locations)
```

### 3. Calculate Emergency Routes

```{r}
# Calculate routes from service locations to accidents
routes <- sero_calculate_routes(
  optimal_locations, 
  data$accident, 
  max_routes = 10
)

print(routes)
```

```{r, fig.cap="Emergency response routes"}
plot(routes)
```

## Complete Analysis Workflow

For convenience, SERO provides a complete analysis function:

```{r}
# Complete analysis with default parameters
analysis <- sero_analyze(
  risk_categories = c(1, 2),
  max_locations = 3,
  max_routes = 10
)

print(analysis)
```

## Interactive Features

SERO includes interactive mapping capabilities (requires plotly):

```{r eval=FALSE}
# Create interactive map for accident selection
interactive_map <- sero_create_interactive_map(data, optimal_locations)
interactive_map

# Select accident by coordinates and calculate route
route <- sero_interactive_routing_workflow(
  data, 
  optimal_locations, 
  longitude = 7.625, 
  latitude = 51.962
)
plot(route)
```

## Advanced Parameter Customization

### Hotspot Analysis Parameters

```{r}
# Fine-tuned hotspot analysis
precise_hotspots <- sero_identify_hotspots(
  data$accident,
  risk_categories = c(1, 2),
  buffer = 500,     # Smaller buffer for precise hotspots
  min_events = 3    # Lower threshold for more hotspots
)

broad_hotspots <- sero_identify_hotspots(
  data$accident,
  risk_categories = c(1, 2),
  buffer = 1500,    # Larger buffer for regional hotspots
  min_events = 8    # Higher threshold for fewer, stronger hotspots
)

cat("Precise hotspots:", nrow(precise_hotspots$hotspots), "\n")
cat("Broad hotspots:", nrow(broad_hotspots$hotspots), "\n")
```

### Multi-Criteria Weight Optimization

```{r}
# Accident-focused optimization
accident_focused <- sero_calculate_optimal_locations(
  data,
  max_locations = 5,
  weight_accidents = 0.6,    # High weight on accident proximity
  weight_population = 0.15,
  weight_roads = 0.1,
  weight_landuse = 0.05,
  weight_distance = 0.05,
  weight_distribution = 0.05
)

# Population-focused optimization
population_focused <- sero_calculate_optimal_locations(
  data,
  max_locations = 5,
  weight_accidents = 0.2,
  weight_population = 0.5,   # High weight on population coverage
  weight_roads = 0.15,
  weight_landuse = 0.05,
  weight_distance = 0.05,
  weight_distribution = 0.05
)

cat("Accident-focused locations:", nrow(accident_focused$locations), "\n")
cat("Population-focused locations:", nrow(population_focused$locations), "\n")
```

## Visualization Functions

SERO provides specialized visualization functions:

```{r, fig.cap="Accident locations by severity"}
sero_plot_accidents(data, color_by_severity = TRUE)
```

```{r, fig.cap="Population density visualization"}
sero_plot_population(data)
```

```{r, fig.cap="Road network visualization"}
sero_plot_roads(data)
```

```{r, fig.cap="Combined spatial analysis visualization"}
sero_plot_combined(data, hotspots, optimal_locations)
```

## Summary

SERO provides a comprehensive toolkit for spatial emergency response optimization:

- **Easy to use**: Simple functions with sensible defaults
- **Flexible**: Extensive parameter customization options
- **Interactive**: Real-time mapping and selection capabilities
- **Well-documented**: Comprehensive examples and documentation
- **Standards-compliant**: Follows R spatial best practices

The package is particularly useful for:

- Urban planners optimizing emergency service placement
- Emergency response coordinators analyzing coverage
- Researchers studying spatial accessibility
- Students learning spatial analysis techniques

For more examples and advanced usage, see the package examples in `inst/examples/`.

## References

- Bivand, R., Pebesma, E., & Gómez-Rubio, V. (2013). Applied spatial data analysis with R. Springer.
- Pebesma, E. (2018). Simple Features for R: Standardized Support for Spatial Vector Data. The R Journal.
- Baddeley, A., Rubak, E., & Turner, R. (2015). Spatial Point Patterns: Methodology and Applications with R. CRC Press.
