---
title: "Spatial Emergency Response Optimization: A Data-Driven Approach to Emergency Service Location Planning"
author: "SERO Development Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Spatial Emergency Response Optimization: A Data-Driven Approach to Emergency Service Location Planning}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 6,
  fig.align = "center"
)
```

# Introduction and Motivation

Emergency response systems face critical challenges in optimally positioning emergency services to minimize response times and maximize coverage. The spatial distribution of accidents exhibits distinct patterns that can inform strategic placement of emergency stations. This analysis demonstrates how spatial statistical methods can identify high-risk areas and optimize emergency service locations using real traffic accident data from Münster, Germany.

The motivation for this work stems from the need to transition from intuition-based to data-driven emergency service planning. Traditional approaches often rely on administrative boundaries or population density alone, potentially missing critical accident concentration areas that emerge from complex urban dynamics. By applying kernel density estimation and spatial optimization algorithms, we can reveal hidden patterns and provide evidence-based recommendations for emergency service deployment.

# Data Description and Exploration

The analysis utilizes a comprehensive spatial dataset from Münster containing traffic accidents, administrative boundaries, road networks, land use patterns, and population distributions. This dataset provides the foundation for understanding spatial accident patterns and optimizing emergency response strategies.

```{r data-loading}
library(SERO)
library(ggplot2)
library(gridExtra)

# Load comprehensive spatial dataset
data <- sero_load_data()

# Examine data structure
cat("Dataset Overview:\n")
cat("- Accidents:", nrow(data$accident), "incidents\n")
cat("- Districts:", nrow(data$districts), "administrative areas\n") 
cat("- Roads:", nrow(data$roads), "network segments\n")
cat("- Population areas:", nrow(data$population), "zones\n")
cat("- Land use areas:", nrow(data$landuse), "polygons\n")
```

```{r accident-severity-analysis}
# Analyze accident severity distribution
severity_counts <- table(data$accident$UKATEGORIE)
severity_df <- data.frame(
  Category = c("Fatal", "Serious", "Light"),
  Count = as.numeric(severity_counts[1:3]),
  Percentage = round(as.numeric(severity_counts[1:3]) / sum(severity_counts) * 100, 1)
)

cat("Accident Severity Distribution:\n")
print(severity_df)

# Visualize severity distribution
ggplot(severity_df, aes(x = Category, y = Count, fill = Category)) +
  geom_col() +
  geom_text(aes(label = paste0(Count, "\n(", Percentage, "%)")), 
            vjust = 1.5, color = "white", fontface = "bold") +
  scale_fill_manual(values = c("Fatal" = "#8B0000", "Serious" = "#FF4500", "Light" = "#FFD700")) +
  labs(title = "Traffic Accident Severity Distribution", 
       subtitle = "Münster Dataset Analysis",
       x = "Accident Severity", y = "Number of Incidents") +
  theme_minimal() +
  theme(legend.position = "none")
```

# Spatial Pattern Analysis

## Hotspot Detection

We identify accident concentration areas using kernel density estimation to reveal spatial clustering patterns invisible to traditional analysis methods.

```{r hotspot-analysis}
# High-risk accident analysis (fatal and serious)
high_risk_accidents <- data$accident[data$accident$UKATEGORIE %in% c(1, 2), ]

cat("High-risk accidents (fatal + serious):", nrow(high_risk_accidents), 
    "of", nrow(data$accident), 
    paste0("(", round(nrow(high_risk_accidents)/nrow(data$accident)*100, 1), "%)"))

# Hotspot identification
hotspots <- sero_hotspots(data$accident, 
                         risk_categories = c(1, 2), 
                         buffer = 1000)

# Display hotspot analysis results
plot(hotspots)

cat("\nHotspot Analysis Results:\n")
cat("- Hotspots identified:", nrow(hotspots$hotspots), "\n")
cat("- Critical areas requiring priority attention\n")
```

## Kernel Density Analysis

Kernel density estimation provides smooth risk surfaces that reveal underlying spatial patterns and inform strategic planning.

```{r heatmap-analysis}
# Create comparative heatmaps for different severity levels
heatmap_combined <- sero_heatmap(data$accident,
                                risk_categories = c(1, 2),
                                bandwidth = 1200,
                                data = data, 
                                color_scheme = "viridis",
                                basemap = "districts") +
  labs(title = "Combined High-Risk Areas", subtitle = "Fatal and serious accidents")

print(heatmap_combined)

# Statistical summary of risk distribution
cat("Risk Surface Analysis:\n")
cat("- Bandwidth used: 1200 meters\n")
cat("- Risk categories: Fatal (1) and Serious (2)\n")
cat("- Analysis reveals concentrated risk zones for targeted intervention\n")
```

# Emergency Service Optimization

## Optimal Location Strategy

We apply spatial optimization algorithms to determine strategic emergency service locations that minimize average response distances to high-risk areas.

```{r optimal-locations}
# Calculate optimal emergency service locations
optimal_3 <- sero_find_optimal_locations(data, 
                                        num_locations = 3,
                                        risk_categories = c(1, 2))

optimal_5 <- sero_find_optimal_locations(data,
                                        num_locations = 5, 
                                        risk_categories = c(1, 2))

optimal_7 <- sero_find_optimal_locations(data,
                                        num_locations = 7,
                                        risk_categories = c(1, 2))

# Performance comparison
performance_data <- data.frame(
  Stations = c(3, 5, 7),
  Avg_Coverage = c(mean(optimal_3$accessibility_score),
                   mean(optimal_5$accessibility_score), 
                   mean(optimal_7$accessibility_score)),
  Max_Distance = c(max(optimal_3$accessibility_score),
                   max(optimal_5$accessibility_score),
                   max(optimal_7$accessibility_score))
)

cat("Optimization Performance Analysis:\n")
print(performance_data)

# Create spatial visualization of optimal locations
library(sf)

# Transform data to consistent CRS for visualization
districts_viz <- st_transform(data$districts, 4326)
landuse_viz <- st_transform(data$landuse, 4326)
roads_viz <- st_transform(data$roads, 4326)
accidents_viz <- st_transform(high_risk_accidents, 4326)

# Transform optimal locations to WGS84
optimal_3_viz <- st_transform(optimal_3, 4326)
optimal_5_viz <- st_transform(optimal_5, 4326)
optimal_7_viz <- st_transform(optimal_7, 4326)

# Create base plot function
create_location_plot <- function(optimal_locations, title, subtitle) {
  ggplot() +
    # Add land use (background)
    geom_sf(data = landuse_viz, fill = "#f0f0f0", color = "#e0e0e0", size = 0.1, alpha = 0.3) +
    # Add roads
    geom_sf(data = roads_viz, color = "#d3d3d3", size = 0.2, alpha = 0.5) +
    # Add district boundaries
    geom_sf(data = districts_viz, fill = "transparent", color = "#2c3e50", size = 0.8) +
    # Add high-risk accidents
    geom_sf(data = accidents_viz, color = "#e74c3c", size = 0.8, alpha = 0.6) +
    # Add optimal locations (top layer)
    geom_sf(data = optimal_locations, size = 4, color = "#2c3e50", fill = "#f39c12", 
            shape = 24, stroke = 1.5) +
    labs(title = title, subtitle = subtitle) +
    theme_void() +
    theme(
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 10, hjust = 0.5),
      plot.margin = margin(5, 5, 5, 5)
    )
}

# Create individual plots
plot_3 <- create_location_plot(optimal_3_viz, "3 Emergency Stations", 
                              paste0("Avg coverage: ", round(mean(optimal_3$accessibility_score), 2)))

plot_5 <- create_location_plot(optimal_5_viz, "5 Emergency Stations", 
                              paste0("Avg coverage: ", round(mean(optimal_5$accessibility_score), 2)))

plot_7 <- create_location_plot(optimal_7_viz, "7 Emergency Stations", 
                              paste0("Avg coverage: ", round(mean(optimal_7$accessibility_score), 2)))

# Performance metrics plot
performance_plot <- ggplot(performance_data, aes(x = Stations)) +
  geom_line(aes(y = Avg_Coverage, color = "Average Coverage"), size = 1.2) +
  geom_line(aes(y = Max_Distance, color = "Maximum Distance"), size = 1.2) +
  geom_point(aes(y = Avg_Coverage, color = "Average Coverage"), size = 3) +
  geom_point(aes(y = Max_Distance, color = "Maximum Distance"), size = 3) +
  labs(title = "Performance Metrics",
       subtitle = "Coverage vs. number of stations",
       x = "Number of Emergency Stations", 
       y = "Distance (accessibility score)",
       color = "Metric") +
  scale_color_manual(values = c("Average Coverage" = "#2E86AB", "Maximum Distance" = "#A23B72")) +
  theme_minimal() +
  theme(plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(size = 10, hjust = 0.5),
        legend.position = "bottom")

# Arrange plots in a 2x2 grid
grid.arrange(plot_3, plot_5, plot_7, performance_plot, ncol = 2, nrow = 2)
```

## Route Analysis

Route optimization calculates emergency response paths from stations to incident locations, providing critical performance metrics.

```{r route-analysis}
# Sample high-risk accidents for route analysis
sample_accidents <- high_risk_accidents[sample(nrow(high_risk_accidents), 15), ]

# Calculate routes using 5-station configuration
routes <- sero_routes(optimal_5, 
                     sample_accidents, 
                     max_routes = 15,
                     data = data)

# Analyze route performance
route_stats <- data.frame(
  Metric = c("Total Routes", "Average Time (min)", "Average Distance (km)", 
             "Routes < 5 min", "Routes 5-10 min", "Routes > 10 min"),
  Value = c(nrow(routes$routes),
            round(mean(routes$routes$estimated_time), 2),
            round(mean(routes$routes$distance)/1000, 2),
            sum(routes$routes$estimated_time < 5),
            sum(routes$routes$estimated_time >= 5 & routes$routes$estimated_time < 10),
            sum(routes$routes$estimated_time >= 10))
)

cat("Route Performance Analysis:\n")
print(route_stats)

# Create spatial visualization of routes
# Transform route data for visualization
routes_viz <- st_transform(routes$routes, 4326)
sample_accidents_viz <- st_transform(sample_accidents, 4326)

# Add time categories to routes
routes_viz$time_category <- cut(routes_viz$estimated_time,
                               breaks = c(0, 5, 10, Inf),
                               labels = c("< 5 min", "5-10 min", "> 10 min"))

# Create route visualization function
create_route_plot <- function(show_landuse = FALSE, show_roads = FALSE, title, subtitle) {
  p <- ggplot()
  
  # Add land use if requested
  if (show_landuse) {
    p <- p + geom_sf(data = landuse_viz, fill = "#f0f0f0", color = "#e0e0e0", 
                     size = 0.1, alpha = 0.3)
  }
  
  # Add roads if requested
  if (show_roads) {
    p <- p + geom_sf(data = roads_viz, color = "#d3d3d3", size = 0.2, alpha = 0.5)
  }
  
  # Add district boundaries
  p <- p + geom_sf(data = districts_viz, fill = "transparent", color = "#2c3e50", size = 0.8)
  
  # Add route lines colored by response time
  p <- p + geom_sf(data = routes_viz, aes(color = time_category), size = 1.2, alpha = 0.8)
  
  # Add sample accidents
  p <- p + geom_sf(data = sample_accidents_viz, color = "#e74c3c", size = 1.5, alpha = 0.8)
  
  # Add emergency stations (top layer)
  p <- p + geom_sf(data = optimal_5_viz, size = 4, color = "#2c3e50", fill = "#f39c12", 
                   shape = 24, stroke = 1.5)
  
  # Add styling and colors
  p <- p + 
    scale_color_manual(values = c("< 5 min" = "#27AE60", "5-10 min" = "#F39C12", "> 10 min" = "#E74C3C"),
                       name = "Response Time") +
    labs(title = title, subtitle = subtitle) +
    theme_void() +
    theme(
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 10, hjust = 0.5),
      legend.position = "bottom",
      legend.title = element_text(size = 10, face = "bold"),
      plot.margin = margin(5, 5, 5, 5)
    )
  
  return(p)
}

# Create different route visualizations
route_basic <- create_route_plot(show_landuse = FALSE, show_roads = FALSE,
                                "Basic Route Network", 
                                "Emergency routes with district boundaries")

route_with_roads <- create_route_plot(show_landuse = FALSE, show_roads = TRUE,
                                     "Routes with Road Network", 
                                     "Infrastructure context for route planning")

route_with_landuse <- create_route_plot(show_landuse = TRUE, show_roads = FALSE,
                                       "Routes with Land Use", 
                                       "Geographic context for emergency response")

# Create performance summary plot
performance_summary <- ggplot(routes_viz) +
  geom_sf(aes(color = time_category), size = 1.5, alpha = 0.8) +
  scale_color_manual(values = c("< 5 min" = "#27AE60", "5-10 min" = "#F39C12", "> 10 min" = "#E74C3C"),
                     name = "Response Time") +
  labs(title = "Route Performance Summary",
       subtitle = paste0("Average time: ", round(mean(routes$routes$estimated_time), 1), " minutes")) +
  theme_void() +
  theme(
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 10, hjust = 0.5),
    legend.position = "bottom",
    legend.title = element_text(size = 10, face = "bold"),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.margin = margin(5, 5, 5, 5)
  ) +
  coord_sf(expand = FALSE)

# Arrange route plots in a 2x2 grid
grid.arrange(route_basic, route_with_roads, route_with_landuse, performance_summary, 
             ncol = 2, nrow = 2)
```

# Discussion of Results

## Key Findings

The spatial analysis reveals several critical insights for emergency service planning in Münster. First, accident hotspots are not uniformly distributed but show distinct clustering patterns, with `r nrow(hotspots$hotspots)` identified high-risk areas requiring priority attention. These hotspots account for a disproportionate share of fatal and serious accidents, representing only `r round(nrow(high_risk_accidents)/nrow(data$accident)*100, 1)`% of total incidents but requiring immediate response capabilities.

The optimization analysis demonstrates clear performance trade-offs between station quantity and coverage efficiency. While 3 stations provide basic coverage with an average accessibility score of `r round(mean(optimal_3$accessibility_score), 2)`, expanding to 5 stations improves coverage to `r round(mean(optimal_5$accessibility_score), 2)` with diminishing returns at 7 stations (`r round(mean(optimal_7$accessibility_score), 2)`). Route analysis confirms that the 5-station configuration achieves `r sum(routes$routes$estimated_time < 5)` routes under 5 minutes, meeting emergency response standards for most high-risk incidents.

## Strategic Recommendations

Based on this analysis, we recommend implementing a 5-station emergency service configuration that balances coverage efficiency with resource constraints. Priority should be given to establishing services near identified hotspots, ensuring rapid response to areas with the highest concentration of severe accidents. The kernel density analysis provides precise coordinates for optimal placement, moving beyond administrative boundaries to data-driven positioning.

Furthermore, the route performance analysis indicates that `r sum(routes$routes$estimated_time >= 10)` routes exceed 10-minute response times, suggesting these areas require additional attention through either supplementary stations or pre-positioned resources during high-risk periods.

## Limitations and Future Work

This analysis focuses on historical accident patterns and assumes static emergency response infrastructure. Future research should incorporate temporal dynamics, traffic patterns, and real-time routing constraints. Additionally, validation against actual emergency response data would strengthen the practical applicability of these recommendations.

# Conclusion

The SERO package demonstrates the value of spatial statistical methods for emergency service optimization. By combining hotspot analysis, kernel density estimation, and optimization algorithms, we can move from intuition-based to evidence-based emergency service planning. The Münster case study shows how these methods reveal hidden patterns in accident data and provide actionable recommendations for improving emergency response coverage and efficiency.
