---
title: "SERO Package: Complete Emergency Response Optimization Workflow"
subtitle: "Hotspots, Heatmaps, Optimal Locations, and Interactive Routing"
author: "SERO Package Demonstration"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: bootstrap
    toc: true
    toc_float: true
    code_folding: show
    fig_width: 12
    fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE, 
  message = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 150
)
```

# 🚨 SERO Package Complete Demonstration

This document demonstrates the complete workflow of the **SERO (Spatial Emergency Response Optimization)** package using real data from Münster, Germany. We'll walk through:

1. **Data Loading** - Import spatial datasets
2. **Hotspot Analysis** - Identify accident clusters using kernel density estimation
3. **Heatmap Visualization** - Create density surfaces with district boundaries
4. **Optimal Location Calculation** - Find best emergency service locations
5. **Route Optimization** - Calculate emergency response routes
6. **Interactive Visualization** - Web-based routing interface

## 📊 Package Setup and Data Loading

```{r libraries}
# Load required packages
library(SERO)
library(sf)
library(ggplot2)
library(dplyr)
library(viridis)
library(scales)

# Check for optional packages
has_leaflet <- requireNamespace("leaflet", quietly = TRUE)
has_osrm <- requireNamespace("osrm", quietly = TRUE)

cat("🚨 SERO Emergency Response Optimization\n")
cat("=====================================\n\n")
cat("Package versions:\n")
cat("- SERO: Available\n")
cat("- sf:", as.character(packageVersion("sf")), "\n")
cat("- ggplot2:", as.character(packageVersion("ggplot2")), "\n")
if (has_leaflet) cat("- leaflet: Available\n")
if (has_osrm) cat("- osrm: Available\n")
```

```{r load_data}
# Load the complete Münster dataset
cat("Loading spatial data for Münster, Germany...\n")
data <- sero_load_data()

cat("Data loaded successfully:\n")
cat("- Road network:", nrow(data$roads), "segments\n")
cat("- Traffic accidents:", nrow(data$accident), "incidents\n")
cat("- Administrative districts:", nrow(data$districts), "areas\n") 
cat("- Land use areas:", nrow(data$landuse), "polygons\n")
cat("- Population density:", nrow(data$population), "areas\n\n")

# Display data extent and coordinate system
cat("Coordinate Reference System:", st_crs(data$accident)$input, "\n")
cat("Study area extent:\n")
bbox <- st_bbox(data$districts)
cat("- West:", round(bbox$xmin, 3), "°\n")
cat("- East:", round(bbox$xmax, 3), "°\n") 
cat("- South:", round(bbox$ymin, 3), "°\n")
cat("- North:", round(bbox$ymax, 3), "°\n")
```

## 🔥 Step 1: Hotspot Analysis

Identify accident hotspots using kernel density estimation, focusing on high-risk incidents.

```{r hotspot_analysis}
cat("Calculating accident hotspots...\n")

# Focus on severe accidents (categories 1 and 2: fatal and serious injury)
high_risk_categories <- c(1, 2)
hotspots <- sero_hotspots(data$accident, risk_categories = high_risk_categories)

cat("Hotspot analysis complete:\n")
cat("- Method: Kernel density estimation\n")
cat("- Risk categories:", paste(high_risk_categories, collapse = ", "), "\n")
cat("- Grid resolution:", attr(hotspots, "grid_size"), "meters\n")
cat("- Bandwidth:", attr(hotspots, "bandwidth"), "meters\n")

# Print summary statistics
print(hotspots)
```

## 🌡️ Step 2: Heatmap Visualization with District Boundaries

Create comprehensive heatmap visualization showing accident density with district boundaries clearly visible.

```{r heatmap_visualization, fig.width=14, fig.height=10}
# Transform data to WGS84 for plotting
districts_wgs84 <- st_transform(data$districts, 4326)
accidents_wgs84 <- st_transform(data$accident, 4326)

# Filter for high-risk accidents for heatmap
high_risk_accidents <- accidents_wgs84[accidents_wgs84$UKATEGORIE %in% high_risk_categories, ]

cat("Creating heatmap visualization...\n")
cat("- Accidents included:", nrow(high_risk_accidents), "high-risk incidents\n")
cat("- Districts:", nrow(districts_wgs84), "administrative areas\n")

# Create comprehensive heatmap
heatmap_plot <- ggplot() +
  # Add district boundaries with semi-transparent fill
  geom_sf(data = districts_wgs84, 
          fill = "white", 
          color = "#2c3e50", 
          size = 1.2, 
          alpha = 0.7) +
  
  # Add accident points with density-based visualization
  stat_density_2d_filled(data = data.frame(st_coordinates(high_risk_accidents)),
                         aes(x = X, y = Y),
                         alpha = 0.6,
                         bins = 15,
                         contour_var = "ndensity") +
  
  # Add individual accident points
  geom_sf(data = high_risk_accidents,
          size = 0.8,
          alpha = 0.4,
          color = "#d73027") +
  
  # Add district labels
  geom_sf_text(data = districts_wgs84,
               aes(label = NAME_STADT),
               size = 3.5,
               fontface = "bold",
               color = "#2c3e50") +
  
  # Styling
  scale_fill_viridis_d(name = "Accident\nDensity", 
                       option = "plasma",
                       alpha = 0.8) +
  
  theme_minimal() +
  theme(
    axis.text = element_text(size = 10),
    legend.position = "right",
    legend.title = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    panel.grid = element_line(color = "grey90", size = 0.5)
  ) +
  
  labs(
    title = "🔥 Emergency Response Hotspot Analysis",
    subtitle = paste("High-Risk Traffic Accidents in Münster Districts |", 
                     nrow(high_risk_accidents), "Fatal & Serious Injury Incidents"),
    x = "Longitude (°)",
    y = "Latitude (°)",
    caption = "Data: OpenStreetMap contributors | Analysis: SERO Package"
  ) +
  
  coord_sf(datum = sf::st_crs(4326))

print(heatmap_plot)
```

## 📍 Step 3: Optimal Location Calculation

Calculate optimal emergency service locations using multi-criteria spatial analysis.

```{r optimal_locations}
cat("Calculating optimal emergency service locations...\n")

# Calculate 6 optimal emergency service locations
optimal_locations <- sero_find_optimal_locations(
  data, 
  num_locations = 6,
  risk_categories = high_risk_categories,
  method = "fast"
)

cat("Optimal location calculation complete:\n")
cat("- Number of stations:", nrow(optimal_locations), "\n")
cat("- Optimization method: Multi-criteria analysis\n")
cat("- Criteria: Accident density, population, road access, land use\n\n")

# Display optimal locations
cat("Optimal Emergency Station Locations:\n")
coords <- st_coordinates(st_transform(optimal_locations, 4326))
for (i in seq_len(nrow(optimal_locations))) {
  cat(sprintf("  Station %d: %.4f°, %.4f° (District: %s)\n", 
              i, coords[i,1], coords[i,2], optimal_locations$NAME_STADT[i]))
}
```

## 🗺️ Step 4: Comprehensive Spatial Visualization

Create a comprehensive map showing hotspots, optimal locations, and district boundaries.

```{r comprehensive_map, fig.width=16, fig.height=12}
# Transform optimal locations to WGS84
optimal_wgs84 <- st_transform(optimal_locations, 4326)
# Add station numbers for labeling
optimal_wgs84$station_id <- 1:nrow(optimal_wgs84)

cat("Creating comprehensive spatial visualization...\n")

# Create comprehensive spatial analysis map
comprehensive_plot <- ggplot() +
  # Base layer: District boundaries
  geom_sf(data = districts_wgs84,
          fill = "white",
          color = "#34495e",
          size = 1.5,
          alpha = 0.8) +
  
  # Add accident density heatmap
  stat_density_2d_filled(data = data.frame(st_coordinates(high_risk_accidents)),
                         aes(x = X, y = Y),
                         alpha = 0.5,
                         bins = 12,
                         contour_var = "ndensity") +
  
  # Add road network (sample for performance)
  geom_sf(data = st_transform(data$roads[sample(nrow(data$roads), 2000), ], 4326),
          color = "grey70",
          size = 0.3,
          alpha = 0.6) +
  
  # Add optimal emergency locations
  geom_sf(data = optimal_wgs84,
          size = 6,
          color = "#2c3e50",
          fill = "#e74c3c",
          shape = 21,
          stroke = 2) +
  
  # Add optimal location labels
  geom_sf_text(data = optimal_wgs84,
               aes(label = paste("Station", station_id)),
               vjust = -1.5,
               size = 3.5,
               fontface = "bold",
               color = "#2c3e50") +
  
  # Add district names
  geom_sf_text(data = districts_wgs84,
               aes(label = NAME_STADT),
               size = 3,
               alpha = 0.8,
               color = "#2c3e50") +
  
  # Styling
  scale_fill_viridis_d(name = "Accident\nDensity",
                       option = "plasma",
                       alpha = 0.7) +
  
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.key.width = unit(1.5, "cm")
  ) +
  
  labs(
    title = "🚨 Complete Emergency Response Optimization Analysis",
    subtitle = "Optimal Station Locations with Accident Hotspots | Münster, Germany",
    caption = "Red circles = Optimal emergency stations | Heatmap = High-risk accident density | Grey lines = Road network"
  )

print(comprehensive_plot)
```

## 🛣️ Step 5: Route Optimization Analysis

Calculate and visualize emergency response routes from optimal locations to accident scenes.

```{r route_analysis}
cat("Calculating emergency response routes...\n")

# Calculate sample routes to demonstrate routing capability
sample_accidents <- data$accident[sample(nrow(data$accident), 8), ]
routes <- sero_routes(optimal_locations, sample_accidents, max_routes = 8, data = data)

cat("Route calculation complete:\n")
print(routes)

# Calculate coverage statistics
if (nrow(routes$routes) > 0) {
  avg_distance <- mean(routes$routes$distance) / 1000  # Use 'distance' column
  avg_time <- mean(routes$routes$estimated_time)
  max_distance <- max(routes$routes$distance) / 1000   # Use 'distance' column  
  max_time <- max(routes$routes$estimated_time)
  
  cat("\nEmergency Response Coverage Analysis:\n")
  cat("- Average response distance:", round(avg_distance, 1), "km\n")
  cat("- Average response time:", round(avg_time, 1), "minutes\n")
  cat("- Maximum response distance:", round(max_distance, 1), "km\n")
  cat("- Maximum response time:", round(max_time, 1), "minutes\n")
}
```

## 📈 Step 6: Route Visualization

Visualize the calculated emergency response routes with spatial context.

```{r route_visualization, fig.width=14, fig.height=10}
if (nrow(routes$routes) > 0) {
  cat("Creating route visualization...\n")
  
  # Transform route data to WGS84
  routes_wgs84 <- st_transform(routes$routes, 4326)
  sample_accidents_wgs84 <- st_transform(sample_accidents[1:nrow(routes$routes), ], 4326)
  
  # Create route visualization
  route_plot <- ggplot() +
    # Base layer: Districts
    geom_sf(data = districts_wgs84,
            fill = "white",
            color = "#34495e",
            size = 1,
            alpha = 0.9) +
    
    # Add road network context
    geom_sf(data = st_transform(data$roads[sample(nrow(data$roads), 1500), ], 4326),
            color = "grey85",
            size = 0.2,
            alpha = 0.7) +
    
    # Add emergency routes
    geom_sf(data = routes_wgs84,
            aes(color = estimated_time),
            size = 2,
            alpha = 0.8) +
    
    # Add optimal emergency stations
    geom_sf(data = optimal_wgs84,
            size = 5,
            color = "#2c3e50",
            fill = "#3498db",
            shape = 21,
            stroke = 2) +
    
    # Add accident locations
    geom_sf(data = sample_accidents_wgs84,
            size = 4,
            color = "#e74c3c",
            shape = 4,
            stroke = 2) +
    
    # Add district labels
    geom_sf_text(data = districts_wgs84,
                 aes(label = NAME_STADT),
                 size = 3,
                 alpha = 0.7,
                 color = "#2c3e50") +
    
    # Color scale for response time
    scale_color_viridis_c(name = "Response\nTime (min)",
                          option = "plasma",
                          trans = "sqrt") +
    
    theme_void() +
    theme(
      legend.position = "bottom",
      legend.title = element_text(size = 12, face = "bold"),
      plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5),
      legend.key.width = unit(1.5, "cm")
    ) +
    
    labs(
      title = "🚑 Emergency Response Route Optimization",
      subtitle = paste("Calculated Routes from Optimal Stations to Accident Locations |", 
                       nrow(routes$routes), "Sample Routes"),
      caption = "Blue circles = Emergency stations | Red X = Accidents | Colored lines = Response routes"
    )
  
  print(route_plot)
} else {
  cat("No routes calculated for visualization.\n")
}
```

## 📊 Step 7: Performance Analytics

Analyze the performance of the emergency response optimization.

```{r performance_analytics, fig.width=12, fig.height=8}
if (nrow(routes$routes) > 0) {
  cat("Creating performance analytics...\n")
  
  # Create performance summary data
  performance_data <- data.frame(
    Route_ID = 1:nrow(routes$routes),
    Distance_km = routes$routes$distance / 1000,  # Use 'distance' column instead of 'distance_m'
    Time_minutes = routes$routes$estimated_time,
    Station_ID = routes$routes$service_location_id,  # Use correct column name
    stringsAsFactors = FALSE
  )
  
  # Distance distribution plot
  dist_plot <- ggplot(performance_data, aes(x = Distance_km)) +
    geom_histogram(bins = 6, fill = "#3498db", alpha = 0.7, color = "white") +
    geom_vline(aes(xintercept = mean(Distance_km)), 
               color = "#e74c3c", linetype = "dashed", size = 1) +
    labs(title = "Response Distance Distribution",
         x = "Distance (km)", y = "Number of Routes") +
    theme_minimal() +
    theme(plot.title = element_text(face = "bold"))
  
  # Time distribution plot  
  time_plot <- ggplot(performance_data, aes(x = Time_minutes)) +
    geom_histogram(bins = 6, fill = "#e74c3c", alpha = 0.7, color = "white") +
    geom_vline(aes(xintercept = mean(Time_minutes)),
               color = "#3498db", linetype = "dashed", size = 1) +
    labs(title = "Response Time Distribution", 
         x = "Time (minutes)", y = "Number of Routes") +
    theme_minimal() +
    theme(plot.title = element_text(face = "bold"))
  
  # Station utilization plot
  station_plot <- ggplot(performance_data, aes(x = factor(Station_ID))) +
    geom_bar(fill = "#2ecc71", alpha = 0.7, color = "white") +
    labs(title = "Station Utilization",
         x = "Station ID", y = "Number of Routes") +
    theme_minimal() +
    theme(plot.title = element_text(face = "bold"))
  
  # Combine plots
  library(gridExtra)
  combined_analytics <- grid.arrange(dist_plot, time_plot, station_plot, 
                                   ncol = 2, nrow = 2,
                                   top = "📊 Emergency Response Performance Analytics")
  
  # Performance summary table
  cat("\nPerformance Summary:\n")
  cat("===================\n")
  summary_stats <- data.frame(
    Metric = c("Average Distance", "Average Time", "Max Distance", "Max Time", 
               "Min Distance", "Min Time", "Total Routes"),
    Value = c(paste(round(mean(performance_data$Distance_km), 1), "km"),
              paste(round(mean(performance_data$Time_minutes), 1), "min"),
              paste(round(max(performance_data$Distance_km), 1), "km"),
              paste(round(max(performance_data$Time_minutes), 1), "min"),
              paste(round(min(performance_data$Distance_km), 1), "km"),
              paste(round(min(performance_data$Time_minutes), 1), "min"),
              nrow(performance_data)),
    stringsAsFactors = FALSE
  )
  
  knitr::kable(summary_stats, caption = "Emergency Response Performance Metrics")
} else {
  cat("No performance data available for analysis.\n")
  cat("Routes object structure:\n")
  if (exists("routes")) {
    print(str(routes))
    if (is.list(routes) && "routes" %in% names(routes)) {
      cat("Routes data columns:", names(routes$routes), "\n")
      cat("Number of routes:", nrow(routes$routes), "\n")
    }
  }
}
```

## 🌐 Step 8: Interactive Routing (Optional)

If the required packages are available, demonstrate interactive routing capability.

```{r interactive_routing, eval=FALSE}
# Note: This chunk is set to eval=FALSE because interactive widgets 
# don't work well in static HTML output. 
# Run this code interactively in R to see the live demo.

if (has_leaflet && has_osrm) {
  cat("Launching interactive routing demonstration...\n")
  cat("Click on the map to simulate emergency incidents!\n")
  
  # Launch the interactive routing demo
  # This will open in your web browser
  source("inst/examples/simple_routing_demo.R")
  
} else {
  cat("Interactive routing requires 'leaflet' and 'osrm' packages.\n")
  cat("Install with: install.packages(c('leaflet', 'osrm'))\n")
}
```

## 📋 Summary and Conclusions

### 🎯 Key Findings

```{r summary}
cat("🚨 SERO Emergency Response Optimization - Summary\n")
cat("================================================\n\n")

cat("📊 Data Analysis:\n")
cat("- Study area: Münster, Germany\n")
cat("- Total accidents analyzed:", nrow(data$accident), "\n")
cat("- High-risk accidents (fatal/serious):", 
    sum(data$accident$UKATEGORIE %in% high_risk_categories), "\n")
cat("- Administrative districts:", nrow(data$districts), "\n\n")

cat("📍 Optimal Locations:\n")
cat("- Emergency stations calculated:", nrow(optimal_locations), "\n")
cat("- Optimization method: Multi-criteria spatial analysis\n")
cat("- Coverage criteria: Accident density, population, accessibility\n\n")

if (nrow(routes$routes) > 0) {
  cat("🛣️ Route Performance:\n")
  cat("- Sample routes calculated:", nrow(routes$routes), "\n")
  cat("- Average response distance:", round(mean(routes$routes$distance)/1000, 1), "km\n")
  cat("- Average response time:", round(mean(routes$routes$estimated_time), 1), "minutes\n")
  cat("- Emergency speed assumption: 30 km/h (urban emergency response)\n\n")
}

cat("🔧 Technical Implementation:\n")
cat("- Hotspot analysis: Kernel density estimation\n") 
cat("- Spatial optimization: Multi-criteria decision analysis\n")
cat("- Route calculation: Road network-based routing\n")
cat("- Visualization: ggplot2 with sf spatial objects\n")
cat("- Interactive features: Shiny + Leaflet (optional)\n\n")

cat("🎯 Applications:\n")
cat("- Emergency service planning and optimization\n")
cat("- Urban planning and public safety analysis\n")
cat("- Research in spatial optimization and GIS\n")
cat("- Educational tool for spatial analysis concepts\n\n")

cat("✅ Analysis complete! All SERO package features demonstrated successfully.\n")
```

### 🚀 Next Steps

- **Extend Analysis**: Try different numbers of emergency stations
- **Custom Data**: Apply SERO to your own city or region  
- **Interactive Exploration**: Run the interactive routing demos
- **Advanced Features**: Explore population weighting and land use constraints
- **Research Applications**: Use SERO for academic or professional projects

### 📞 Resources

- **Package Documentation**: `help(package = "SERO")`
- **Interactive Examples**: See `inst/examples/` directory
- **Source Code**: Available on GitHub
- **Support**: Report issues and feature requests

---

**🎉 Congratulations!** You've successfully explored the complete SERO emergency response optimization workflow. The package provides powerful tools for spatial analysis, optimization, and visualization of emergency service systems.

```{r session_info}
cat("Session Information:\n")
cat("===================\n")
sessionInfo()
```
