# Diagnostic: check min and max before any code runs
```{r}
```
title: "SERO Emergency Response Optimization: Example Workflow"
author: "SERO Package Team"
date: "2025-07-13"
output:
  pdf_document:
    toc: true
    number_sections: true
---

# Setup: load libraries and restore base min/max
```{r setup, include=TRUE}
library(SERO)
library(sf)
library(ggplot2)
library(MASS)
library(raster)
# Restore base min/max functions AFTER loading all packages
min <- base::min
max <- base::max
cat("[DIAG] Final check - min is base::min:", identical(min, base::min), "\n")
cat("[DIAG] Final check - max is base::max:", identical(max, base::max), "\n")
---
# Introduction

This document demonstrates a typical user workflow for emergency location optimization using the SERO package. It covers data loading, accident visualization, optimal location finding, and professional mapping, with all plots and results shown inline.

# 1. Load Data

```{r}
library(SERO)
library(sf)
library(ggplot2)
# Load demo data (replace with your own data as needed)
data <- sero_load_data()
str(data)
```

# 2. Visualize Accidents by Severity

```{r}
cat("\n[DIAG] Structure of data$accident before plotting:\n"); str(data$accident)
cat("[DIAG] Type of data$accident:", typeof(data$accident), "\n")
cat("[DIAG] Structure of data$districts before plotting:\n"); str(data$districts)
cat("[DIAG] Type of data$districts:", typeof(data$districts), "\n")
accident_plot <- sero_plot_accidents(
  accidents = data$accident,
  districts = data$districts
)
accident_plot
```

# 3. Find Optimal Emergency Locations

```{r}
cat("\n[DIAG] Structure of data before finding optimal locations:\n"); str(data)
cat("[DIAG] Type of data:", typeof(data), "\n")
optimal_locs <- sero_find_optimal_locations(
  data = data,
  num_locations = 5,
  method = "fast",
  risk_category_names = c("fatal", "serious"),
  quick = TRUE
)
cat("[DIAG] Structure of optimal_locs after finding:\n"); str(optimal_locs)
cat("[DIAG] Type of optimal_locs:", typeof(optimal_locs), "\n")
print(optimal_locs)
```



```{r}
cat("\n[DIAG] Structure of optimal_locs before plotting:\n"); str(optimal_locs)
cat("[DIAG] Type of optimal_locs:", typeof(optimal_locs), "\n")
cat("[DIAG] Structure of data$districts before plotting optimal locations:\n"); str(data$districts)
cat("[DIAG] Type of data$districts:", typeof(data$districts), "\n")
optimal_plot <- sero_plot_optimal_locations(
  districts = data$districts,
  roads = data$roads,
  landuse = data$landuse,
  accidents = data$accident,
  district_border_color = "black",
  road_border_color = "gray",
  landuse_border_color = "darkgreen",
  landuse_fill_color = "lightgreen",
  landuse_sample_size = 100,
  road_sample_size = 200,
  optimal_color = "red",
  optimal_shape = 17
)
optimal_plot
```



# 5. Create Accident Hotspots (Kernel Density)

```{r}
cat("\n[DIAG] Structure of data$accident before hotspot plot:\n"); str(data$accident)
cat("[DIAG] Type of data$accident:", typeof(data$accident), "\n")
hotspot_plot <- sero_hotspots(data$accident, districts = data$districts, bandwidth = 0.01)
hotspot_plot
```

# 6. Create Accident Heatmap (2D Histogram)

```{r}
cat("\n[DIAG] Structure of data$accident before heatmap plot:\n"); str(data$accident)
cat("[DIAG] Type of data$accident:", typeof(data$accident), "\n")
heatmap_plot <- sero_heatmap(data$accident, districts = data$districts, bins = 50)
heatmap_plot
```


# 7. Performance Summary

```{r}
cat("\n[DIAG] Structure of optimal_locs before summary:\n"); str(optimal_locs)
cat("[DIAG] Type of optimal_locs:", typeof(optimal_locs), "\n")
cat("Emergency bases found:", nrow(optimal_locs), "\n")
if ("accessibility_score" %in% names(optimal_locs)) {
  cat("Average accessibility score:", round(mean(optimal_locs$accessibility_score), 3), "\n")
  cat("Best accessibility score:", round(max(optimal_locs$accessibility_score), 3), "\n")
}
if ("accident_count_500m" %in% names(optimal_locs)) {
  cat("Average accidents covered (500m):", round(mean(optimal_locs$accident_count_500m), 1), "\n")
  cat("Maximum accidents covered (500m):", max(optimal_locs$accident_count_500m), "\n")
}
```


# 8. Save Results

```{r}
sero_save_optimal_locations(optimal_locs, "optimal_locations.sqlite")
```


# 9. Session Info

```{r}
sessionInfo()
```
