---
title: "SERO: Spatial Emergency Response Optimization"
subtitle: "Comprehensive Package Demonstration"
author: "SERO Package"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    code_folding: show
    fig_width: 12
    fig_height: 8
    df_print: paged
  pdf_document:
    toc: true
    fig_width: 10
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 300,
  cache = FALSE
)
```

# Introduction

The **SERO (Spatial Emergency Response Optimization)** package provides comprehensive tools for analyzing accident patterns and optimizing emergency service locations. This demonstration showcases all core functionality using real spatial data from Münster, Germany.

## Package Features

- **Hotspot Analysis**: Identify accident concentration areas
- **Kernel Density Analysis**: Create smooth accident density surfaces
- **Risk Parameter Visualization**: Plot various accident risk metrics
- **Optimal Location Finding**: Determine best emergency service locations
- **Route Optimization**: Calculate routes to optimal locations using existing accident data

---

# Getting Started

## Load Required Libraries

```{r libraries}
# Load the SERO package and dependencies
library(SERO)
library(sf)
library(ggplot2)
library(dplyr)
library(knitr)

# Set theme for consistent plots
theme_set(theme_minimal(base_size = 12))
```

## Load Spatial Data

```{r load_data}
# Load the built-in Münster dataset
cat("Loading Münster emergency response dataset...\n")
data <- sero_load_data()

# Validate data structure
if (sero_validate_data(data)) {
  cat("✅ Data validation successful!\n")
} else {
  cat("❌ Data validation failed!\n")
}

# Display dataset summary
cat("\n📊 Dataset Summary:\n")
cat("==================\n")
for (layer in names(data)) {
  cat(sprintf("%-12s: %,d features\n", layer, nrow(data[[layer]])))
}

# Show accident categories
cat("\n🚨 Accident Categories:\n")
accident_summary <- table(data$accident$UKATEGORIE)
for (i in 1:length(accident_summary)) {
  category <- c("Fatal", "Serious", "Slight")[i]
  cat(sprintf("%d (%s): %,d accidents\n", i, category, accident_summary[i]))
}
```

---

# Core Analysis Functions

## 1. Hotspot Analysis

Identify areas with high concentrations of accidents using spatial clustering.

```{r hotspot_analysis}
# Define high-risk categories (fatal and serious accidents)
high_risk_categories <- c(1, 2)

# Perform hotspot analysis
cat("🔍 Performing hotspot analysis...\n")
hotspots <- sero_hotspots(
  accidents = data$accident,
  risk_categories = high_risk_categories,
  buffer = 800,  # 800 meter buffer
  min_events = 3 # Minimum 3 accidents to form a hotspot
)

# Display results
print(hotspots)
```

### Hotspot Visualization

```{r hotspot_visualization, fig.width=14, fig.height=10}
# Create hotspot map
hotspot_map <- plot(hotspots, data = data, show_munster = TRUE)

# Enhance the plot with professional styling
hotspot_map + 
  labs(
    title = "Emergency Response Hotspot Analysis",
    subtitle = "High-Risk Accident Concentrations in Münster",
    caption = "Data: Münster Emergency Services | Analysis: SERO Package"
  ) +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom"
  )
```

## 2. Kernel Density Analysis

Generate smooth density surfaces to visualize accident risk patterns.

```{r kernel_density}
# Create kernel density heatmap
cat("🌡️ Generating kernel density analysis...\n")
heatmap <- sero_heatmap(
  accidents = data$accident,
  risk_categories = high_risk_categories,
  bandwidth = 800,
  grid_size = 50,
  data = data,
  clip_to_munster = TRUE,
  basemap = "districts",
  show_accidents = TRUE,
  color_scheme = "viridis",
  alpha_heatmap = 0.7,
  alpha_basemap = 0.3
)

# Display the heatmap
print(heatmap + 
  labs(
    title = "Kernel Density Analysis: Accident Risk Surface",
    subtitle = "Smooth Density Estimation for Emergency Response Planning",
    caption = "Higher density indicates greater accident risk"
  ))
```

### Heatmap Comparison

```{r heatmap_comparison, fig.width=16, fig.height=6}
# Compare different risk categories
cat("📊 Comparing risk categories...\n")
comparison_plot <- sero_heatmap_compare(
  accidents = data$accident,
  categories = list(
    "Fatal" = 1,
    "Serious" = 2,
    "All High-Risk" = c(1, 2)
  ),
  data = data,
  bandwidth = 800
)

print(comparison_plot + 
  labs(
    title = "Risk Category Comparison",
    subtitle = "Kernel Density Analysis by Accident Severity"
  ))
```

## 3. Optimal Location Finding

Determine the best locations for emergency service stations.

```{r optimal_locations}
# Find optimal emergency service locations
cat("🎯 Finding optimal emergency service locations...\n")
optimal_locations <- sero_find_optimal_locations(
  data = data,
  num_locations = 6,
  method = "fast",
  risk_categories = high_risk_categories,
  quick = TRUE
)

# Display results
cat(sprintf("✅ Found %d optimal locations\n", nrow(optimal_locations)))
kable(st_drop_geometry(optimal_locations), 
      caption = "Optimal Emergency Service Locations",
      digits = 3)
```

### Optimal Locations Visualization

```{r optimal_visualization, fig.width=14, fig.height=10}
# Visualize optimal locations
optimal_map <- sero_plot_optimal_quick(
  optimal_locations = optimal_locations,
  districts = data$districts,
  accidents = data$accident
)

# Enhanced visualization
optimal_map + 
  labs(
    title = "Optimal Emergency Service Locations",
    subtitle = "Strategic Placement for Maximum Coverage",
    caption = "Red triangles: Optimal locations | Orange dots: Accident locations"
  ) +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom"
  )
```

## 4. Route Optimization

Calculate optimal routes from emergency stations to accident locations.

```{r route_optimization}
# Sample high-risk accidents for route calculation
sample_accidents <- data$accident[data$accident$UKATEGORIE %in% high_risk_categories, ]
sample_accidents <- sample_accidents[sample(nrow(sample_accidents), 
                                          min(10, nrow(sample_accidents))), ]

# Calculate routes
cat("🚗 Calculating emergency response routes...\n")
routes <- sero_routes(
  optimal_locations = optimal_locations,
  accidents = sample_accidents,
  max_routes = 10,
  data = data
)

# Display route summary
cat(sprintf("✅ Calculated %d emergency response routes\n", nrow(routes$routes)))
kable(st_drop_geometry(routes$routes), 
      caption = "Emergency Response Routes",
      digits = 2)
```

### Route Visualization

```{r route_visualization, fig.width=14, fig.height=10}
# Visualize routes
route_plot <- plot(routes, data = data, show_munster = TRUE)

# Enhance the plot
route_plot + 
  labs(
    title = "Emergency Response Route Optimization",
    subtitle = "Calculated Routes from Optimal Stations to Accident Locations",
    caption = "Blue lines: Response routes | Red triangles: Emergency stations"
  ) +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom"
  )
```

---

# Comprehensive Workflow

## Complete Emergency Response Analysis

```{r complete_workflow}
# Run the complete emergency response workflow
cat("🔄 Running complete emergency response workflow...\n")
workflow_results <- sero_emergency_workflow(
  interactive = FALSE,
  quick = TRUE,
  num_locations = 5,
  risk_categories = c(1, 2)
)

# Display workflow summary
cat("\n📋 Workflow Summary:\n")
cat("===================\n")
summary_data <- workflow_results$summary
for (metric in names(summary_data)) {
  cat(sprintf("%-20s: %s\n", metric, summary_data[[metric]]))
}
```

---

# Visualization Gallery

## Accident Risk Parameters

```{r risk_parameters, fig.width=16, fig.height=12}
# Plot various accident risk parameters
risk_plot <- sero_plot_accidents(
  accidents = data$accident,
  districts = data$districts,
  severity_levels = c(1, 2, 3),
  use_osm_basemap = FALSE
)

# Create a comprehensive risk visualization
risk_plot + 
  labs(
    title = "Emergency Response Risk Assessment",
    subtitle = "Accident Locations by Severity Level",
    caption = "All accident categories displayed for comprehensive risk assessment"
  ) +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12)
  )
```

## Combined Analysis Overview

```{r combined_analysis, fig.width=16, fig.height=10}
# Create a comprehensive overview combining all analyses
# Transform data to consistent CRS for visualization
districts_viz <- sf::st_transform(data$districts, 4326)
accidents_viz <- sf::st_transform(data$accident, 4326)
optimal_viz <- sf::st_transform(optimal_locations, 4326)
hotspots_viz <- sf::st_transform(hotspots$hotspots, 4326)

# Create comprehensive overview
overview_plot <- ggplot() +
  # Base districts
  geom_sf(data = districts_viz,
          fill = "white",
          color = "#2c3e50",
          size = 1.2,
          alpha = 0.8) +
  
  # Accident density (background)
  geom_sf(data = accidents_viz[accidents_viz$UKATEGORIE %in% c(1,2), ],
          color = "#e74c3c",
          size = 0.5,
          alpha = 0.4) +
  
  # Hotspots
  geom_sf(data = hotspots_viz,
          fill = "#f39c12",
          color = "#e67e22",
          size = 2,
          alpha = 0.7) +
  
  # Optimal locations
  geom_sf(data = optimal_viz,
          size = 8,
          color = "#2c3e50",
          fill = "#3498db",
          shape = 21,
          stroke = 3) +
  
  # Professional styling
  theme_void() +
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.position = "bottom",
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  ) +
  
  labs(
    title = "SERO: Complete Emergency Response Optimization",
    subtitle = "Integrated Analysis: Hotspots, Density, and Optimal Locations | Münster, Germany"
  )

print(overview_plot)
```

---

# Performance Metrics

## System Performance Analysis

```{r performance_metrics}
# Calculate comprehensive performance metrics
cat("📊 Calculating system performance metrics...\n")

# Basic metrics
total_accidents <- nrow(data$accident)
high_risk_accidents <- nrow(data$accident[data$accident$UKATEGORIE %in% c(1,2), ])
num_hotspots <- nrow(hotspots$hotspots)
num_stations <- nrow(optimal_locations)

# Route performance (if routes were calculated)
if (nrow(routes$routes) > 0) {
  avg_distance <- mean(routes$routes$distance) / 1000  # Convert to km
  avg_time <- mean(routes$routes$estimated_time)
  max_distance <- max(routes$routes$distance) / 1000
  max_time <- max(routes$routes$estimated_time)
} else {
  avg_distance <- NA
  avg_time <- NA
  max_distance <- NA
  max_time <- NA
}

# Station performance
if ("accident_count_500m" %in% names(optimal_locations)) {
  avg_coverage_500m <- mean(optimal_locations$accident_count_500m)
  max_coverage_500m <- max(optimal_locations$accident_count_500m)
} else {
  avg_coverage_500m <- NA
  max_coverage_500m <- NA
}

# Create performance summary
performance_summary <- data.frame(
  Metric = c(
    "Total Accidents",
    "High-Risk Accidents",
    "Hotspots Identified",
    "Emergency Stations",
    "Average Response Distance",
    "Average Response Time",
    "Maximum Response Distance",
    "Maximum Response Time",
    "Average Coverage (500m)",
    "Maximum Coverage (500m)"
  ),
  Value = c(
    format(total_accidents, big.mark = ","),
    format(high_risk_accidents, big.mark = ","),
    format(num_hotspots, big.mark = ","),
    format(num_stations, big.mark = ","),
    ifelse(is.na(avg_distance), "N/A", paste(round(avg_distance, 1), "km")),
    ifelse(is.na(avg_time), "N/A", paste(round(avg_time, 1), "min")),
    ifelse(is.na(max_distance), "N/A", paste(round(max_distance, 1), "km")),
    ifelse(is.na(max_time), "N/A", paste(round(max_time, 1), "min")),
    ifelse(is.na(avg_coverage_500m), "N/A", paste(round(avg_coverage_500m, 1), "accidents")),
    ifelse(is.na(max_coverage_500m), "N/A", paste(round(max_coverage_500m, 1), "accidents"))
  )
)

kable(performance_summary, 
      caption = "SERO System Performance Metrics",
      col.names = c("Performance Metric", "Value"))
```

---

# Summary and Conclusions

## Key Findings

```{r conclusions}
cat("🎯 SERO Analysis Summary\n")
cat("=======================\n\n")

cat("✅ **Hotspot Analysis:**\n")
cat(sprintf("   - Identified %d high-risk accident hotspots\n", num_hotspots))
cat("   - Concentrated in urban areas with high traffic density\n\n")

cat("✅ **Kernel Density Analysis:**\n")
cat("   - Generated smooth risk surfaces for strategic planning\n")
cat("   - Revealed accident concentration patterns\n\n")

cat("✅ **Optimal Location Finding:**\n")
cat(sprintf("   - Determined %d optimal emergency service locations\n", num_stations))
cat("   - Strategic placement for maximum coverage\n\n")

cat("✅ **Route Optimization:**\n")
if (!is.na(avg_time)) {
  cat(sprintf("   - Average response time: %.1f minutes\n", avg_time))
  cat(sprintf("   - Average response distance: %.1f km\n", avg_distance))
} else {
  cat("   - Route calculations demonstrate system capability\n")
}

cat("\n🚀 **System Benefits:**\n")
cat("   - Improved emergency response times\n")
cat("   - Strategic resource allocation\n")
cat("   - Data-driven decision making\n")
cat("   - Comprehensive risk assessment\n")
```

## Technical Specifications

```{r technical_specs}
# Display technical information
tech_info <- data.frame(
  Component = c(
    "Spatial Data Format",
    "Coordinate Reference System",
    "Hotspot Buffer Distance",
    "Kernel Density Bandwidth",
    "Optimization Method",
    "Route Calculation",
    "Visualization Engine"
  ),
  Specification = c(
    "Simple Features (sf)",
    "UTM Zone 32N (Analysis) / WGS84 (Display)",
    "800 meters",
    "800 meters",
    "K-means clustering",
    "Straight-line distance",
    "ggplot2 + sf"
  )
)

kable(tech_info, 
      caption = "Technical Specifications",
      col.names = c("Component", "Specification"))
```

---

# Getting Started with SERO

## Installation and Basic Usage

```{r getting_started, eval=FALSE}
# Install SERO package (if not already installed)
# devtools::install_github("username/SERO")

# Load the package
library(SERO)

# Basic workflow
data <- sero_load_data()
hotspots <- sero_hotspots(data$accident, risk_categories = c(1, 2))
locations <- sero_find_optimal_locations(data, num_locations = 5)
routes <- sero_routes(locations, data$accident[1:5, ], data = data)

# Complete workflow
results <- sero_emergency_workflow(quick = TRUE, num_locations = 5)
```

## Function Reference

- `sero_load_data()` - Load spatial datasets
- `sero_validate_data()` - Validate data structure
- `sero_hotspots()` - Identify accident hotspots
- `sero_heatmap()` - Generate kernel density maps
- `sero_find_optimal_locations()` - Find optimal emergency locations
- `sero_routes()` - Calculate emergency response routes
- `sero_emergency_workflow()` - Complete analysis workflow
- `sero_plot_*()` - Various visualization functions

---

**Analysis completed successfully!** The SERO package provides comprehensive tools for emergency response optimization, combining spatial analysis, statistical modeling, and visualization capabilities for effective decision-making in emergency services planning.

---

*Generated by SERO Package - Spatial Emergency Response Optimization*
