---
title: "SERO: Spatial Emergency Response Optimization"
subtitle: "Comprehensive Package Demonstration"
author: "SERO Package"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    code_folding: show
    fig_width: 12
    fig_height: 8
    fig_caption: true
vignette: >
  %\VignetteIndexEntry{SERO Package Demonstration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 12,
  fig.height = 8,
  fig.align = "center",
  cache = FALSE,
  results = 'markup'
)
```

# Overview

The **SERO** (Spatial Emergency Response Optimization) package provides comprehensive tools for analyzing and optimizing emergency response systems using spatial data. This demonstration showcases the core functions and their applications in emergency service planning.

## Key Features

- **Spatial Data Analysis**: Load and analyze emergency service spatial datasets
- **Hotspot Detection**: Identify high-risk accident concentration areas
- **Heatmap Visualization**: Create kernel density maps for risk assessment
- **Location Optimization**: Find optimal emergency service locations
- **Route Planning**: Calculate fastest emergency response routes
- **Basemap Integration**: Visualize results with multiple spatial contexts

```{r libraries}
# Load required libraries
library(SERO)
library(sf)
library(ggplot2)
library(gridExtra)
library(grid)  # For textGrob function
library(dplyr)
library(knitr)

# Set plotting theme
theme_set(theme_minimal(base_size = 10))
```

# Data Loading and Validation

```{r data-loading}
# Load the built-in Münster dataset
cat("Loading Münster emergency response dataset...\n")
data <- sero_load_data()

# Validate data structure
if (sero_validate_data(data)) {
  cat("✅ Data validation successful!\n")
} else {
  cat("❌ Data validation failed!\n")
}

# Display dataset summary
cat("\n📊 Dataset Summary:\n")
cat("==================\n")
for (layer in names(data)) {
  cat(sprintf("%-12s: %d features\n", layer, nrow(data[[layer]])))
}

# Show accident categories
cat("\n🚨 Accident Categories:\n")
accident_summary <- table(data$accident$UKATEGORIE)
for (i in seq_along(accident_summary)) {
  category <- c("Fatal", "Serious", "Slight")[i]
  cat(sprintf("%d (%s): %d accidents\n", i, category, accident_summary[i]))
}
```

# Temporal Analysis

Understanding when accidents occur helps optimize emergency response scheduling and resource allocation.

```{r temporal-analysis, fig.cap="Temporal distribution of accidents showing seasonal and time patterns"}
# Extract temporal information from accidents
accidents_temporal <- data$accident

# Create time-based analysis if temporal columns exist
if ("UZEIT" %in% names(accidents_temporal) || "hour" %in% names(accidents_temporal)) {
  
  # Extract hour information (assuming UZEIT or similar time column exists)
  if ("UZEIT" %in% names(accidents_temporal)) {
    accidents_temporal$hour <- as.numeric(substr(accidents_temporal$UZEIT, 1, 2))
  }
  
  # Create hourly distribution
  hourly_dist <- accidents_temporal %>%
    st_drop_geometry() %>%
    filter(!is.na(hour)) %>%
    group_by(hour, UKATEGORIE) %>%
    summarise(count = n(), .groups = 'drop') %>%
    mutate(category = case_when(
      UKATEGORIE == 1 ~ "Fatal",
      UKATEGORIE == 2 ~ "Serious", 
      UKATEGORIE == 3 ~ "Slight"
    ))
  
  p1 <- ggplot(hourly_dist, aes(x = hour, y = count, fill = category)) +
    geom_bar(stat = "identity", alpha = 0.8) +
    scale_fill_manual(values = c("Fatal" = "#d32f2f", "Serious" = "#f57c00", "Slight" = "#388e3c")) +
    labs(title = "Accidents by Hour of Day",
         subtitle = "Emergency response demand patterns",
         x = "Hour of Day", y = "Number of Accidents", fill = "Severity") +
    theme_minimal() +
    theme(plot.title = element_text(size = 12, face = "bold"))
  
} else {
  # Create a simple severity distribution if no time data
  severity_dist <- accidents_temporal %>%
    st_drop_geometry() %>%
    group_by(UKATEGORIE) %>%
    summarise(count = n(), .groups = 'drop') %>%
    mutate(category = case_when(
      UKATEGORIE == 1 ~ "Fatal",
      UKATEGORIE == 2 ~ "Serious", 
      UKATEGORIE == 3 ~ "Slight"
    ))
  
  p1 <- ggplot(severity_dist, aes(x = category, y = count, fill = category)) +
    geom_bar(stat = "identity", alpha = 0.8) +
    scale_fill_manual(values = c("Fatal" = "#d32f2f", "Serious" = "#f57c00", "Slight" = "#388e3c")) +
    labs(title = "Accident Distribution by Severity",
         subtitle = "Emergency response demand by severity level",
         x = "Accident Severity", y = "Number of Accidents", fill = "Severity") +
    theme_minimal() +
    theme(plot.title = element_text(size = 12, face = "bold"),
          legend.position = "none")
}

# Create spatial distribution by district
if ("districts" %in% names(data)) {
  # Spatial join to get accidents per district with CRS alignment
  accidents_aligned <- st_transform(data$accident, st_crs(data$districts))
  accidents_by_district <- st_join(accidents_aligned, data$districts) %>%
    st_drop_geometry() %>%
    group_by(UKATEGORIE) %>%
    summarise(count = n(), .groups = 'drop') %>%
    mutate(category = case_when(
      UKATEGORIE == 1 ~ "Fatal",
      UKATEGORIE == 2 ~ "Serious", 
      UKATEGORIE == 3 ~ "Slight"
    ))
  
  p2 <- ggplot(accidents_by_district, aes(x = reorder(category, -count), y = count, fill = category)) +
    geom_bar(stat = "identity", alpha = 0.8) +
    scale_fill_manual(values = c("Fatal" = "#d32f2f", "Serious" = "#f57c00", "Slight" = "#388e3c")) +
    labs(title = "High-Risk Categories Distribution",
         subtitle = "Focus on fatal and serious accidents for emergency planning",
         x = "Accident Category", y = "Number of Accidents", fill = "Severity") +
    theme_minimal() +
    theme(plot.title = element_text(size = 12, face = "bold"),
          legend.position = "none")
} else {
  p2 <- p1  # Fallback to same plot
}

# Combine temporal plots
grid.arrange(p1, p2, ncol = 2, nrow = 1)
```

# Hotspot Analysis

Hotspot analysis identifies areas with high concentrations of emergency incidents, helping prioritize resource allocation.

```{r hotspot-analysis}
# Define high-risk categories (fatal and serious accidents)
high_risk_categories <- c(1, 2)

# Perform hotspot analysis with different parameters
cat("🔍 Performing hotspot analysis...\n")

# Standard hotspot analysis
hotspots_standard <- sero_hotspots(
  accidents = data$accident,
  risk_categories = high_risk_categories,
  buffer = 800,  # 800 meter buffer
  min_events = 3 # Minimum 3 accidents to form a hotspot
)

# Sensitive hotspot analysis (smaller buffer, fewer events)
hotspots_sensitive <- sero_hotspots(
  accidents = data$accident,
  risk_categories = high_risk_categories,
  buffer = 500,  # 500 meter buffer
  min_events = 2 # Minimum 2 accidents to form a hotspot
)

# Conservative hotspot analysis (larger buffer, more events)
hotspots_conservative <- sero_hotspots(
  accidents = data$accident,
  risk_categories = high_risk_categories,
  buffer = 1000, # 1000 meter buffer
  min_events = 5  # Minimum 5 accidents to form a hotspot
)

# Display results
cat("Standard hotspots found:", nrow(hotspots_standard$hotspots), "\n")
cat("Sensitive hotspots found:", nrow(hotspots_sensitive$hotspots), "\n")
cat("Conservative hotspots found:", nrow(hotspots_conservative$hotspots), "\n")
```

## Hotspot Comparison Visualization

```{r hotspot-comparison, fig.cap="Comparison of hotspot detection methods with different sensitivity levels"}
# Create hotspot comparison plots with consistent styling and shared legend
p1 <- plot(hotspots_standard, data = data, show_munster = TRUE) +
  labs(title = "Standard Hotspots",
       subtitle = "800m buffer, min 3 events") +
  theme(plot.title = element_text(size = 12, face = "bold"),
        legend.position = "none")

p2 <- plot(hotspots_sensitive, data = data, show_munster = TRUE) +
  labs(title = "Sensitive Hotspots",
       subtitle = "500m buffer, min 2 events") +
  theme(plot.title = element_text(size = 12, face = "bold"),
        legend.position = "none")

p3 <- plot(hotspots_conservative, data = data, show_munster = TRUE) +
  labs(title = "Conservative Hotspots",
       subtitle = "1000m buffer, min 5 events") +
  theme(plot.title = element_text(size = 12, face = "bold"),
        legend.position = "none")

# Create shared legend plot
p4 <- plot(hotspots_standard, data = data, show_munster = TRUE) +
  labs(title = "Hotspot Analysis Legend",
       subtitle = "Severity Score = Fatal×3 + Serious×2") +
  theme(plot.title = element_text(size = 12, face = "bold"),
        legend.position = "bottom",
        legend.box = "horizontal",
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  guides(
    fill = guide_colorbar(title = "Severity Score", title.position = "top"),
    size = guide_legend(title = "Accident Count", title.position = "top"),
    color = guide_legend(title = "Category", title.position = "top")
  )

# Combine plots with shared legend
grid.arrange(p1, p2, p3, p4, ncol = 2, nrow = 2)
```

# Heatmap Analysis

Heatmap visualization provides continuous risk surfaces using kernel density estimation.

```{r heatmap-analysis}
# Create heatmaps with different parameters
cat("🌡️ Generating heatmap analysis...\n")

# Standard heatmap
heatmap_standard <- sero_heatmap(
  accidents = data$accident,
  risk_categories = high_risk_categories,
  bandwidth = 800,
  grid_size = 50,
  data = data,
  clip_to_munster = TRUE,
  basemap = "districts",
  show_accidents = TRUE,
  color_scheme = "viridis",
  alpha_heatmap = 0.7,
  alpha_basemap = 0.3
)

# High resolution heatmap
heatmap_hires <- sero_heatmap(
  accidents = data$accident,
  risk_categories = high_risk_categories,
  bandwidth = 500,  # Smaller bandwidth for more detail
  grid_size = 75,   # Higher resolution
  data = data,
  clip_to_munster = TRUE,
  basemap = "districts",
  show_accidents = FALSE,
  color_scheme = "plasma",
  alpha_heatmap = 0.8,
  alpha_basemap = 0.2
)

# Smooth heatmap (larger bandwidth)
heatmap_smooth <- sero_heatmap(
  accidents = data$accident,
  risk_categories = high_risk_categories,
  bandwidth = 1200, # Larger bandwidth for smoother surface
  grid_size = 40,   # Lower resolution
  data = data,
  clip_to_munster = TRUE,
  basemap = "districts",
  show_accidents = FALSE,
  color_scheme = "inferno",
  alpha_heatmap = 0.6,
  alpha_basemap = 0.4
)
```

## Heatmap Comparison Visualization

```{r heatmap-comparison, fig.cap="Comparison of kernel density heatmaps with different bandwidth parameters"}
# Enhance heatmap plots with consistent titles
heatmap_standard_plot <- heatmap_standard + 
  labs(title = "Standard Resolution Heatmap",
       subtitle = "Bandwidth: 800m, Grid: 50x50") +
  theme(plot.title = element_text(size = 12, face = "bold"))

heatmap_hires_plot <- heatmap_hires + 
  labs(title = "High Resolution Heatmap",
       subtitle = "Bandwidth: 500m, Grid: 75x75") +
  theme(plot.title = element_text(size = 12, face = "bold"))

heatmap_smooth_plot <- heatmap_smooth + 
  labs(title = "Smooth Heatmap",
       subtitle = "Bandwidth: 1200m, Grid: 40x40") +
  theme(plot.title = element_text(size = 12, face = "bold"))

# Create category comparison  
comparison_plot <- sero_heatmap_compare(
  accidents = data$accident,
  data = data,
  analysis_types = c("urban", "regional", "overview")
)

# Combine plots (showing first 3 from comparison)
if(length(comparison_plot) >= 3) {
  grid.arrange(heatmap_standard_plot, heatmap_hires_plot, 
               heatmap_smooth_plot, comparison_plot[[1]], 
               ncol = 2, nrow = 2)
} else {
  # Fallback with simple comparison
  simple_comparison <- sero_heatmap(
    accidents = data$accident,
    risk_categories = c(1, 2),
    bandwidth = 800,
    grid_size = 50,
    data = data,
    clip_to_munster = TRUE,
    basemap = "districts",
    show_accidents = FALSE,
    color_scheme = "magma",
    alpha_heatmap = 0.7,
    alpha_basemap = 0.3
  ) +
    labs(title = "Combined Risk Categories",
         subtitle = "Fatal and serious accidents combined") +
    theme(plot.title = element_text(size = 12, face = "bold"))
  
  grid.arrange(heatmap_standard_plot, heatmap_hires_plot, 
               heatmap_smooth_plot, simple_comparison, 
               ncol = 2, nrow = 2)
}
```

# Optimal Location Analysis

Finding optimal emergency service locations is crucial for minimizing response times.

```{r optimal-locations}
# Find optimal locations with different approaches
cat("🎯 Finding optimal emergency service locations...\n")

# Standard optimization (3 locations)
optimal_3 <- sero_find_optimal_locations(
  data = data,
  num_locations = 3,
  method = "fast",
  risk_categories = high_risk_categories
)

# More locations (5 locations)
optimal_5 <- sero_find_optimal_locations(
  data = data,
  num_locations = 5,
  method = "fast",
  risk_categories = high_risk_categories
)

# Extensive coverage (7 locations)
optimal_7 <- sero_find_optimal_locations(
  data = data,
  num_locations = 7,
  method = "fast",
  risk_categories = high_risk_categories
)

# Display results
cat("✅ Optimal locations found:\n")
cat("  3 locations: Average coverage =", round(mean(optimal_3$accessibility_score), 2), "\n")
cat("  5 locations: Average coverage =", round(mean(optimal_5$accessibility_score), 2), "\n")
cat("  7 locations: Average coverage =", round(mean(optimal_7$accessibility_score), 2), "\n")
```

## Optimal Location Comparison

```{r optimal-comparison, fig.cap="Comparison of optimal emergency service locations with different coverage strategies"}
# Create optimal location plots
p1 <- sero_plot_optimal_quick(optimal_3, data$districts, accidents = data$accident) +
  labs(title = "3 Emergency Stations",
       subtitle = paste("Avg coverage:", round(mean(optimal_3$accessibility_score), 2))) +
  theme(plot.title = element_text(size = 12, face = "bold"))

p2 <- sero_plot_optimal_quick(optimal_5, data$districts, accidents = data$accident) +
  labs(title = "5 Emergency Stations",
       subtitle = paste("Avg coverage:", round(mean(optimal_5$accessibility_score), 2))) +
  theme(plot.title = element_text(size = 12, face = "bold"))

p3 <- sero_plot_optimal_quick(optimal_7, data$districts, accidents = data$accident) +
  labs(title = "7 Emergency Stations",
       subtitle = paste("Avg coverage:", round(mean(optimal_7$accessibility_score), 2))) +
  theme(plot.title = element_text(size = 12, face = "bold"))

# Create coverage analysis plot
coverage_data <- data.frame(
  locations = c(3, 5, 7),
  avg_coverage = c(mean(optimal_3$accessibility_score),
                   mean(optimal_5$accessibility_score),
                   mean(optimal_7$accessibility_score))
)

p4 <- ggplot(coverage_data, aes(x = locations, y = avg_coverage)) +
  geom_line(color = "blue", size = 1.2) +
  geom_point(color = "red", size = 3) +
  labs(title = "Coverage vs Number of Stations",
       subtitle = "Relationship between stations and coverage score",
       x = "Number of Emergency Stations",
       y = "Average Coverage Score") +
  theme(plot.title = element_text(size = 12, face = "bold"))

# Combine plots
grid.arrange(p1, p2, p3, p4, ncol = 2, nrow = 2)
```

# Route Optimization and Basemap Integration

Route optimization calculates the fastest paths from emergency stations to incident locations.

```{r route-analysis}
# Create routes with different scenarios
cat("🚗 Calculating emergency response routes...\n")

# Use optimal 5-location setup for routing
sample_accidents <- data$accident[data$accident$UKATEGORIE %in% high_risk_categories, ]
sample_accidents <- sample_accidents[sample(nrow(sample_accidents), 15), ]

# Create routes
routes <- sero_routes(
  locations = optimal_5,
  accidents = sample_accidents,
  max_routes = 15,
  data = data
)

# Display route summary
cat("✅ Route Analysis Results:\n")
cat("  Total routes calculated:", nrow(routes$routes), "\n")
cat("  Average travel time:", round(mean(routes$routes$estimated_time), 2), "minutes\n")
cat("  Average distance:", round(mean(routes$routes$distance)/1000, 2), "km\n")
cat("  Routes under 5 minutes:", sum(routes$routes$estimated_time < 5), "\n")
cat("  Routes 5-10 minutes:", sum(routes$routes$estimated_time >= 5 & routes$routes$estimated_time < 10), "\n")
cat("  Routes over 10 minutes:", sum(routes$routes$estimated_time >= 10), "\n")
```

## Route Visualization with Different Basemaps

```{r route-basemaps, fig.cap="Emergency response routes displayed with different spatial contexts"}
# Create enhanced route visualization function
enhanced_route_plot <- function(routes, optimal_locations, sample_accidents, data, 
                               show_roads = TRUE, show_landuse = TRUE, show_districts = TRUE) {
  
  # Check if routes object has the expected structure
  if (is.null(routes$routes) || nrow(routes$routes) == 0) {
    # Create a basic plot with locations and accidents
    p <- ggplot()
    
    # Add district boundaries
    if (show_districts && "districts" %in% names(data)) {
      districts_viz <- sf::st_transform(data$districts, 4326)
      p <- p + geom_sf(data = districts_viz,
                       fill = "transparent",
                       color = "#2c3e50",
                       size = 1.2,
                       alpha = 0.8)
    }
    
    # Add accident locations
    accidents_viz <- sf::st_transform(sample_accidents, 4326)
    p <- p + geom_sf(data = accidents_viz,
                     color = "#e74c3c",
                     size = 2,
                     alpha = 0.7)
    
    # Add emergency stations (top layer)
    locations_viz <- sf::st_transform(optimal_locations, 4326)
    p <- p + geom_sf(data = locations_viz,
                     size = 6,
                     color = "#2c3e50",
                     fill = "#f39c12",
                     shape = 24,
                     stroke = 2)
    
    p <- p + 
      theme_void() +
      theme(
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(size = 10, hjust = 0.5),
        legend.position = "none",
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA),
        plot.margin = margin(10, 10, 10, 10)
      )
    
    return(p)
  }
  
  # Transform data to consistent CRS for visualization
  routes_viz <- sf::st_transform(routes$routes, 4326)
  accidents_viz <- sf::st_transform(sample_accidents, 4326)
  locations_viz <- sf::st_transform(optimal_locations, 4326)
  
  # Base plot
  p <- ggplot()
  
  # Add landuse layer (background)
  if (show_landuse && "landuse" %in% names(data)) {
    landuse_viz <- sf::st_transform(data$landuse, 4326)
    p <- p + geom_sf(data = landuse_viz,
                     fill = "#f5f5f5",
                     color = "#e0e0e0",
                     size = 0.1,
                     alpha = 0.3)
  }
  
  # Add road network
  if (show_roads && "roads" %in% names(data)) {
    roads_viz <- sf::st_transform(data$roads, 4326)
    p <- p + geom_sf(data = roads_viz,
                     color = "#d3d3d3",
                     size = 0.2,
                     alpha = 0.4)
  }
  
  # Add district boundaries
  if (show_districts && "districts" %in% names(data)) {
    districts_viz <- sf::st_transform(data$districts, 4326)
    p <- p + geom_sf(data = districts_viz,
                     fill = "transparent",
                     color = "#2c3e50",
                     size = 1.2,
                     alpha = 0.8)
  }
  
  # Add route lines
  p <- p + geom_sf(data = routes_viz,
                   color = "#3498db",
                   size = 1.5,
                   alpha = 0.8)
  
  # Add accident locations
  p <- p + geom_sf(data = accidents_viz,
                   color = "#e74c3c",
                   size = 2,
                   alpha = 0.7)
  
  # Add emergency stations (TOP LAYER - always visible)
  p <- p + geom_sf(data = locations_viz,
                   size = 8,
                   color = "#2c3e50",
                   fill = "#f39c12",
                   shape = 24,
                   stroke = 2)
  
  # Professional styling optimized for grid layout
  p <- p + 
    theme_void() +
    theme(
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 10, hjust = 0.5),
      legend.position = "none",
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA),
      plot.margin = margin(10, 10, 10, 10)
    )
  
  return(p)
}

# Create four versions of the route visualization for comprehensive grid display
route_basic <- enhanced_route_plot(routes, optimal_5, sample_accidents, data, 
                                  show_roads = FALSE, show_landuse = FALSE, show_districts = TRUE) +
  labs(title = "Basic Route Analysis", 
       subtitle = "Districts and Emergency Stations")

route_with_roads <- enhanced_route_plot(routes, optimal_5, sample_accidents, data, 
                                       show_roads = TRUE, show_landuse = FALSE, show_districts = TRUE) +
  labs(title = "Routes with Road Network", 
       subtitle = "Including Transportation Infrastructure")

route_with_landuse <- enhanced_route_plot(routes, optimal_5, sample_accidents, data, 
                                         show_roads = FALSE, show_landuse = TRUE, show_districts = TRUE) +
  labs(title = "Routes with Land Use", 
       subtitle = "Geographic Context and Zoning")

route_comprehensive <- enhanced_route_plot(routes, optimal_5, sample_accidents, data, 
                                         show_roads = TRUE, show_landuse = TRUE, show_districts = TRUE) +
  labs(title = "Comprehensive Analysis", 
       subtitle = "Complete Geographic Context")

# Create a 2x2 grid layout for route visualizations
route_grid <- grid.arrange(
  route_basic,
  route_with_roads,
  route_with_landuse,
  route_comprehensive,
  ncol = 2,
  nrow = 2,
  top = textGrob("Emergency Response Route Optimization: Multi-Context Analysis", 
                 gp = gpar(fontsize = 16, fontface = "bold")),
  bottom = textGrob("Progressive visualization from basic districts to comprehensive geographic context", 
                    gp = gpar(fontsize = 11, fontface = "italic"))
)

route_grid
```

## Route Performance Analysis

```{r route-performance, fig.cap="Analysis of emergency response route performance and accessibility"}
# Create performance analysis
routes_sf <- routes$routes
routes_sf$time_category <- cut(routes_sf$estimated_time,
                              breaks = c(0, 5, 10, Inf),
                              labels = c("< 5 min", "5-10 min", "> 10 min"))

# Performance summary plot
performance_summary <- routes_sf %>%
  st_drop_geometry() %>%
  group_by(time_category) %>%
  summarise(count = n(), .groups = 'drop')

p1 <- ggplot(performance_summary, aes(x = time_category, y = count, fill = time_category)) +
  geom_bar(stat = "identity", alpha = 0.8) +
  scale_fill_manual(values = c("green", "orange", "red"), guide = "none") +
  labs(title = "Route Performance Distribution",
       subtitle = "Number of routes by response time category",
       x = "Response Time Category",
       y = "Number of Routes") +
  theme(plot.title = element_text(size = 12, face = "bold"))

# Distance vs Time analysis
p2 <- ggplot(st_drop_geometry(routes_sf), aes(x = distance/1000, y = estimated_time)) +
  geom_point(aes(color = time_category), size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "black", linetype = "dashed") +
  scale_color_manual(values = c("green", "orange", "red"), name = "Time Category") +
  labs(title = "Distance vs Travel Time",
       subtitle = "Relationship between route distance and time",
       x = "Distance (km)",
       y = "Travel Time (minutes)") +
  theme(plot.title = element_text(size = 12, face = "bold"))

# Service area coverage
service_coverage <- routes$service_locations %>%
  st_drop_geometry() %>%
  group_by(location_id) %>%
  summarise(
    accidents_500m = first(accident_count_500m),
    accidents_1000m = first(accident_count_1000m),
    .groups = 'drop'
  )

p3 <- ggplot(service_coverage, aes(x = location_id)) +
  geom_bar(aes(y = accidents_500m), stat = "identity", fill = "lightblue", alpha = 0.7) +
  geom_bar(aes(y = accidents_1000m), stat = "identity", fill = "darkblue", alpha = 0.5) +
  labs(title = "Service Area Coverage",
       subtitle = "Accidents within 500m (light) and 1000m (dark) of stations",
       x = "Emergency Station",
       y = "Number of Accidents") +
  theme(plot.title = element_text(size = 12, face = "bold"))

# Accessibility heatmap
routes_wgs84 <- st_transform(routes_sf, 4326)
locations_wgs84 <- st_transform(routes$service_locations, 4326)
districts_wgs84 <- st_transform(data$districts, 4326)

p4 <- ggplot() +
  geom_sf(data = districts_wgs84, fill = "lightgray", color = "darkgray", alpha = 0.3) +
  geom_sf(data = routes_wgs84, aes(color = estimated_time), size = 1.2, alpha = 0.8) +
  geom_sf(data = locations_wgs84, color = "red", size = 4, shape = 3, stroke = 1.5) +
  scale_color_viridis_c(name = "Travel Time\n(minutes)", option = "plasma") +
  theme_void() +
  labs(title = "Emergency Response Accessibility",
       subtitle = "Spatial distribution of response times") +
  theme(plot.title = element_text(size = 12, face = "bold"),
        legend.position = "bottom")

# Combine plots
grid.arrange(p1, p2, p3, p4, ncol = 2, nrow = 2)
```

# Complete Workflow Integration

The SERO package provides an integrated workflow for comprehensive emergency response analysis.

```{r complete-workflow}
# Run complete emergency response workflow
cat("🔄 Running complete emergency response workflow...\n")

workflow_results <- sero_emergency_workflow(
  interactive = FALSE,
  quick = TRUE,
  num_locations = 4,
  risk_categories = c(1, 2)
)

# Display workflow summary
cat("📋 Complete Workflow Results:\n")
cat("==============================\n")
summary_data <- workflow_results$summary
for (metric in names(summary_data)) {
  cat(sprintf("%-25s: %s\n", metric, summary_data[[metric]]))
}
```

# Performance Metrics Summary

```{r performance-summary}
# Create comprehensive performance summary
cat("📊 SERO Package Performance Summary\n")
cat("===================================\n")

# Data statistics
total_accidents <- nrow(data$accident)
high_risk_accidents <- nrow(data$accident[data$accident$UKATEGORIE %in% c(1,2), ])
districts_count <- nrow(data$districts)

# Hotspot statistics
hotspot_counts <- c(
  standard = nrow(hotspots_standard$hotspots),
  sensitive = nrow(hotspots_sensitive$hotspots),
  conservative = nrow(hotspots_conservative$hotspots)
)

# Location optimization statistics
location_coverage <- c(
  three_locations = mean(optimal_3$accessibility_score),
  five_locations = mean(optimal_5$accessibility_score),
  seven_locations = mean(optimal_7$accessibility_score)
)

# Route performance statistics
route_performance <- c(
  total_routes = nrow(routes$routes),
  avg_time = mean(routes$routes$estimated_time),
  avg_distance = mean(routes$routes$distance)/1000,
  fast_routes = sum(routes$routes$estimated_time < 5),
  medium_routes = sum(routes$routes$estimated_time >= 5 & routes$routes$estimated_time < 10),
  slow_routes = sum(routes$routes$estimated_time >= 10)
)

# Create summary table
summary_table <- data.frame(
  Metric = c("Total Accidents", "High-Risk Accidents", "Districts",
             "Standard Hotspots", "Sensitive Hotspots", "Conservative Hotspots",
             "3-Location Coverage", "5-Location Coverage", "7-Location Coverage",
             "Total Routes", "Avg Travel Time (min)", "Avg Distance (km)",
             "Fast Routes (<5min)", "Medium Routes (5-10min)", "Slow Routes (>10min)"),
  Value = c(total_accidents, high_risk_accidents, districts_count,
            hotspot_counts,
            round(location_coverage, 2),
            route_performance[1], round(route_performance[2:3], 2),
            route_performance[4:6])
)

kable(summary_table, caption = "SERO Package Analysis Summary", align = "lr")
```

# Key Findings and Recommendations

## Analysis Results

1. **Hotspot Analysis**: Identified varying numbers of high-risk areas depending on sensitivity parameters
   - Standard approach: `r nrow(hotspots_standard$hotspots)` hotspots
   - Sensitive approach: `r nrow(hotspots_sensitive$hotspots)` hotspots  
   - Conservative approach: `r nrow(hotspots_conservative$hotspots)` hotspots

2. **Location Optimization**: Coverage improves with more emergency stations
   - 3 stations: `r round(mean(optimal_3$accessibility_score), 2)` coverage score
   - 5 stations: `r round(mean(optimal_5$accessibility_score), 2)` coverage score
   - 7 stations: `r round(mean(optimal_7$accessibility_score), 2)` coverage score

3. **Route Performance**: Most routes achieve acceptable response times
   - `r sum(routes$routes$estimated_time < 5)` routes under 5 minutes
   - `r sum(routes$routes$estimated_time >= 5 & routes$routes$estimated_time < 10)` routes 5-10 minutes
   - `r sum(routes$routes$estimated_time >= 10)` routes over 10 minutes

## Recommendations

1. **Optimal Configuration**: Use 5 emergency stations for balanced coverage and efficiency
2. **Hotspot Focus**: Prioritize standard hotspot detection for resource allocation
3. **Route Monitoring**: Monitor routes over 10 minutes for improvement opportunities
4. **Basemap Integration**: Use comprehensive basemap view for strategic planning

# Conclusion

The SERO package provides a comprehensive toolkit for emergency response optimization, offering:

- **Flexible Analysis**: Multiple approaches for hotspot detection and location optimization
- **Visual Integration**: Rich basemap support for spatial context
- **Performance Metrics**: Detailed analysis of response times and coverage
- **Workflow Integration**: Complete end-to-end analysis capabilities

This demonstration shows how SERO can be used to make data-driven decisions for emergency service planning and optimization.

---

*Generated with SERO v`r packageVersion("SERO")` on `r Sys.Date()`*
