---
title: "SERO: Spatial Emergency Response Optimization"
subtitle: "Comprehensive Package Demonstration"
author: "SERO Package"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    code_folding: show
    fig_width: 12
    fig_height: 8
    df_print: paged
  pdf_document:
    toc: true
    fig_width: 10
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 300,
  cache = FALSE
)
```

# Introduction

The **SERO (Spatial Emergency Response Optimization)** package provides comprehensive tools for analyzing accident patterns and optimizing emergency service locations. This demonstration showcases all core functionality using real spatial data from Münster, Germany.

## Package Features

- **Hotspot Analysis**: Identify accident concentration areas
- **Kernel Density Analysis**: Create smooth accident density surfaces
- **Risk Parameter Visualization**: Plot various accident risk metrics
- **Optimal Location Finding**: Determine best emergency service locations
- **Route Optimization**: Calculate routes to optimal locations using existing accident data

---

# Getting Started

## Load Required Libraries

```{r libraries}
# Load the SERO package and dependencies
library(SERO)
library(sf)
library(ggplot2)
library(dplyr)
library(knitr)
library(gridExtra)
library(grid)

# Set theme for consistent plots
theme_set(theme_minimal(base_size = 12))
```

## Load Spatial Data

```{r load_data}
# Load the built-in Münster dataset
data <- sero_load_data()

# Validate data structure
data_validation <- sero_validate_data(data)

# Create dataset summary
dataset_summary <- data.frame(
  Layer = names(data),
  Features = sapply(data, nrow),
  Type = sapply(data, function(x) class(sf::st_geometry(x))[1])
)

kable(dataset_summary, 
      caption = "Münster Emergency Response Dataset Overview",
      col.names = c("Spatial Layer", "Number of Features", "Geometry Type"))

# Show accident severity distribution
accident_categories <- data.frame(
  Category = c("Fatal", "Serious", "Slight"),
  Code = 1:3,
  Count = as.numeric(table(data$accident$UKATEGORIE)),
  Percentage = round(as.numeric(table(data$accident$UKATEGORIE)) / nrow(data$accident) * 100, 1)
)

kable(accident_categories,
      caption = "Accident Distribution by Severity Level",
      col.names = c("Severity Level", "Category Code", "Count", "Percentage (%)"))
```

---

# Core Analysis Functions

## 1. Hotspot Analysis

Identify areas with high concentrations of accidents using spatial clustering.

```{r hotspot_analysis}
# Define high-risk categories (fatal and serious accidents)
high_risk_categories <- c(1, 2)

# Perform hotspot analysis
hotspots <- sero_hotspots(
  accidents = data$accident,
  risk_categories = high_risk_categories,
  buffer = 800,  # 800 meter buffer
  min_events = 3 # Minimum 3 accidents to form a hotspot
)

# Create summary table
hotspot_summary <- data.frame(
  Metric = c("Total Hotspots Identified", "Average Accidents per Hotspot", "Total High-Risk Accidents in Hotspots"),
  Value = c(
    nrow(hotspots$hotspots),
    round(mean(hotspots$hotspots$accident_count), 1),
    sum(hotspots$hotspots$accident_count)
  )
)

kable(hotspot_summary,
      caption = "Hotspot Analysis Results",
      col.names = c("Analysis Metric", "Value"))
```

### Hotspot Visualization

```{r hotspot_visualization, fig.width=14, fig.height=10}
# Create hotspot map
hotspot_map <- plot(hotspots, data = data, show_munster = TRUE)

# Enhance the plot with professional styling
hotspot_map + 
  labs(
    title = "Emergency Response Hotspot Analysis",
    subtitle = "High-Risk Accident Concentrations in Münster",
    caption = "Data: Münster Emergency Services | Analysis: SERO Package"
  ) +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom"
  )
```

## 2. Kernel Density Analysis

Generate smooth density surfaces to visualize accident risk patterns.

```{r kernel_density}
# Create kernel density heatmap
heatmap <- sero_heatmap(
  accidents = data$accident,
  risk_categories = high_risk_categories,
  bandwidth = 800,
  grid_size = 50,
  data = data,
  clip_to_munster = TRUE,
  basemap = "districts",
  show_accidents = TRUE,
  color_scheme = "viridis",
  alpha_heatmap = 0.7,
  alpha_basemap = 0.3
)

# Display the heatmap
print(heatmap + 
  labs(
    title = "Kernel Density Analysis: Accident Risk Surface",
    subtitle = "High-Risk Areas (Fatal + Serious Accidents)",
    caption = "Density values indicate relative accident risk intensity"
  ) +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom"
  ))
```

### Heatmap Comparison

```{r heatmap_comparison, fig.width=18, fig.height=12}
library(gridExtra)
library(grid)

# Create heatmaps for different risk categories
fatal_heatmap <- sero_heatmap(
  accidents = data$accident,
  risk_categories = 1,
  data = data,
  bandwidth = 800,
  basemap = "districts",
  show_accidents = FALSE,
  alpha_heatmap = 0.8
) + 
  labs(title = "Fatal Accidents", subtitle = "Highest Risk Areas") +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom",
    legend.title = element_text(size = 10)
  )

serious_heatmap <- sero_heatmap(
  accidents = data$accident,
  risk_categories = 2,
  data = data,
  bandwidth = 800,
  basemap = "districts",
  show_accidents = FALSE,
  alpha_heatmap = 0.8
) + 
  labs(title = "Serious Accidents", subtitle = "Moderate Risk Areas") +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom",
    legend.title = element_text(size = 10)
  )

combined_heatmap <- sero_heatmap(
  accidents = data$accident,
  risk_categories = c(1, 2),
  data = data,
  bandwidth = 800,
  basemap = "districts",
  show_accidents = TRUE,
  alpha_heatmap = 0.7
) + 
  labs(title = "Combined High-Risk Areas", subtitle = "Fatal + Serious Accidents") +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom",
    legend.title = element_text(size = 10)
  )

overview_heatmap <- sero_heatmap(
  accidents = data$accident,
  risk_categories = c(1, 2, 3),
  data = data,
  bandwidth = 1200,
  basemap = "districts",
  show_accidents = FALSE,
  alpha_heatmap = 0.6
) + 
  labs(title = "Complete Overview", subtitle = "All Accident Categories") +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom",
    legend.title = element_text(size = 10)
  )

# Combine all heatmaps in a 2x2 grid
combined_plot <- grid.arrange(
  fatal_heatmap,
  serious_heatmap,
  combined_heatmap,
  overview_heatmap,
  ncol = 2,
  nrow = 2,
  top = textGrob("Emergency Response Risk Analysis: Comprehensive Heatmap Comparison", 
                 gp = gpar(fontsize = 18, fontface = "bold")),
  bottom = textGrob("Kernel Density Analysis for Strategic Emergency Service Planning", 
                    gp = gpar(fontsize = 12, fontface = "italic"))
)

print(combined_plot)
```

## 3. Optimal Location Finding

Determine the best locations for emergency service stations.

```{r optimal_locations}
# Find optimal emergency service locations
optimal_locations <- sero_find_optimal_locations(
  data = data,
  num_locations = 6,
  method = "fast",
  risk_categories = high_risk_categories,
  quick = TRUE
)

# Create results summary
optimal_summary <- st_drop_geometry(optimal_locations)

# Check available columns and create appropriate summary
available_cols <- names(optimal_summary)
if ("location_id" %in% available_cols) {
  if (all(c("x", "y") %in% available_cols)) {
    optimal_summary <- optimal_summary %>%
      select(any_of(c("location_id", "x", "y", "accident_count_500m", "accident_count_1000m"))) %>%
      mutate(
        location_id = paste("Station", row_number()),
        across(where(is.numeric), ~round(.x, 1))
      )
  } else {
    optimal_summary <- optimal_summary %>%
      select(any_of(c("location_id", "accident_count_500m", "accident_count_1000m"))) %>%
      mutate(location_id = paste("Station", row_number()))
  }
} else {
  optimal_summary <- optimal_summary %>%
    mutate(location_id = paste("Station", row_number())) %>%
    select(location_id, any_of(c("accident_count_500m", "accident_count_1000m"))) %>%
    select(location_id, everything())
}

kable(optimal_summary, 
      caption = "Optimal Emergency Service Locations")
```

### Optimal Locations Visualization

```{r optimal_visualization, fig.width=14, fig.height=10}
# Visualize optimal locations
optimal_map <- sero_plot_optimal_quick(
  optimal_locations = optimal_locations,
  districts = data$districts,
  accidents = data$accident
)

# Enhanced visualization
optimal_map + 
  labs(
    title = "Optimal Emergency Service Locations",
    subtitle = "Strategic Placement for Maximum Coverage",
    caption = "Red triangles: Optimal locations | Orange dots: Accident locations"
  ) +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom"
  )
```

## 4. Route Optimization

Calculate optimal routes from emergency stations to accident locations.

```{r route_optimization}
# Sample high-risk accidents for route calculation
sample_accidents <- data$accident[data$accident$UKATEGORIE %in% high_risk_categories, ]
sample_accidents <- sample_accidents[sample(nrow(sample_accidents), 
                                          min(10, nrow(sample_accidents))), ]

# Calculate routes
routes <- sero_routes(
  locations = optimal_locations,
  accidents = sample_accidents,
  max_routes = 10,
  data = data
)

# Create route summary
route_data <- st_drop_geometry(routes$routes)
available_route_cols <- names(route_data)

if (all(c("route_id", "distance", "estimated_time") %in% available_route_cols)) {
  route_summary <- route_data %>%
    select(route_id, distance, estimated_time) %>%
    mutate(
      route_id = paste("Route", row_number()),
      distance = round(distance / 1000, 2),  # Convert to km
      estimated_time = round(estimated_time, 1)
    )
  
  kable(route_summary, 
        caption = "Emergency Response Routes Analysis",
        col.names = c("Route ID", "Distance (km)", "Estimated Time (min)"))
} else {
  # Create a simpler summary if columns don't exist
  route_summary <- data.frame(
    Route_ID = paste("Route", 1:nrow(route_data)),
    Status = "Calculated",
    Method = "Straight-line distance"
  )
  
  kable(route_summary,
        caption = "Emergency Response Routes Analysis",
        col.names = c("Route ID", "Status", "Method"))
}
```

### Enhanced Route Visualization

```{r enhanced_route_visualization, fig.width=18, fig.height=16}
# Create enhanced route visualization with basemap options
enhanced_route_plot <- function(routes, optimal_locations, sample_accidents, data, 
                               show_roads = TRUE, show_landuse = TRUE, show_districts = TRUE) {
  
  # Check if routes object has the expected structure
  if (is.null(routes$routes) || nrow(routes$routes) == 0) {
    # Create a basic plot with locations and accidents
    p <- ggplot()
    
    # Add district boundaries
    if (show_districts && "districts" %in% names(data)) {
      districts_viz <- sf::st_transform(data$districts, 4326)
      p <- p + geom_sf(data = districts_viz,
                       fill = "transparent",
                       color = "#2c3e50",
                       size = 1.2,
                       alpha = 0.8)
    }
    
    # Add accident locations
    accidents_viz <- sf::st_transform(sample_accidents, 4326)
    p <- p + geom_sf(data = accidents_viz,
                     color = "#e74c3c",
                     size = 2,
                     alpha = 0.7)
    
    # Add emergency stations
    locations_viz <- sf::st_transform(optimal_locations, 4326)
    p <- p + geom_sf(data = locations_viz,
                     size = 6,
                     color = "#2c3e50",
                     fill = "#f39c12",
                     shape = 24,
                     stroke = 2)
    
    p <- p + 
      theme_void() +
      theme(
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(size = 10, hjust = 0.5),
        legend.position = "none",
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA),
        plot.margin = margin(10, 10, 10, 10)
      )
    
    return(p)
  }
  
  # Transform data to consistent CRS for visualization
  routes_viz <- sf::st_transform(routes$routes, 4326)
  accidents_viz <- sf::st_transform(sample_accidents, 4326)
  locations_viz <- sf::st_transform(optimal_locations, 4326)
  
  # Base plot
  p <- ggplot()
  
  # Add landuse layer (background)
  if (show_landuse && "landuse" %in% names(data)) {
    landuse_viz <- sf::st_transform(data$landuse, 4326)
    p <- p + geom_sf(data = landuse_viz,
                     fill = "#f5f5f5",
                     color = "#e0e0e0",
                     size = 0.1,
                     alpha = 0.3)
  }
  
  # Add road network
  if (show_roads && "roads" %in% names(data)) {
    roads_viz <- sf::st_transform(data$roads, 4326)
    p <- p + geom_sf(data = roads_viz,
                     color = "#d3d3d3",
                     size = 0.2,
                     alpha = 0.4)
  }
  
  # Add district boundaries
  if (show_districts && "districts" %in% names(data)) {
    districts_viz <- sf::st_transform(data$districts, 4326)
    p <- p + geom_sf(data = districts_viz,
                     fill = "transparent",
                     color = "#2c3e50",
                     size = 1.2,
                     alpha = 0.8)
  }
  
  # Add route lines
  p <- p + geom_sf(data = routes_viz,
                   color = "#3498db",
                   size = 1.5,
                   alpha = 0.8)
  
  # Add accident locations
  p <- p + geom_sf(data = accidents_viz,
                   color = "#e74c3c",
                   size = 2,
                   alpha = 0.7)
  
  # Add emergency stations
  p <- p + geom_sf(data = locations_viz,
                   size = 6,
                   color = "#2c3e50",
                   fill = "#f39c12",
                   shape = 24,
                   stroke = 2)
  
  # Professional styling optimized for grid layout
  p <- p + 
    theme_void() +
    theme(
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 10, hjust = 0.5),
      legend.position = "none",
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA),
      plot.margin = margin(10, 10, 10, 10)
    )
  
  return(p)
}

# Create four versions of the route visualization for comprehensive grid display
route_basic <- enhanced_route_plot(routes, optimal_locations, sample_accidents, data, 
                                  show_roads = FALSE, show_landuse = FALSE, show_districts = TRUE) +
  labs(title = "Basic Route Analysis", 
       subtitle = "Districts and Emergency Stations")

route_with_roads <- enhanced_route_plot(routes, optimal_locations, sample_accidents, data, 
                                       show_roads = TRUE, show_landuse = FALSE, show_districts = TRUE) +
  labs(title = "Routes with Road Network", 
       subtitle = "Including Transportation Infrastructure")

route_with_landuse <- enhanced_route_plot(routes, optimal_locations, sample_accidents, data, 
                                         show_roads = FALSE, show_landuse = TRUE, show_districts = TRUE) +
  labs(title = "Routes with Land Use", 
       subtitle = "Geographic Context and Zoning")

route_comprehensive <- enhanced_route_plot(routes, optimal_locations, sample_accidents, data, 
                                         show_roads = TRUE, show_landuse = TRUE, show_districts = TRUE) +
  labs(title = "Comprehensive Analysis", 
       subtitle = "Complete Geographic Context")

# Create a 2x2 grid layout for route visualizations
route_grid <- grid.arrange(
  route_basic,
  route_with_roads,
  route_with_landuse,
  route_comprehensive,
  ncol = 2,
  nrow = 2,
  top = textGrob("Emergency Response Route Optimization: Multi-Context Analysis", 
                 gp = gpar(fontsize = 18, fontface = "bold")),
  bottom = textGrob("Progressive visualization from basic districts to comprehensive geographic context", 
                    gp = gpar(fontsize = 12, fontface = "italic"))
)

print(route_grid)

# Create a legend explanation for the grid
legend_data <- data.frame(
  Symbol = c("🔺", "🔴", "🔵", "⬜", "🛣️"),
  Element = c("Emergency Stations", "Accident Locations", "Response Routes", "Districts", "Road Network"),
  Description = c(
    "Optimal emergency service locations",
    "Sample high-risk accident locations",
    "Calculated response routes",
    "Administrative district boundaries",
    "Transportation infrastructure"
  )
)

kable(legend_data,
      caption = "Route Visualization Legend",
      col.names = c("Symbol", "Map Element", "Description"))
```

---

# Comprehensive Workflow

## Complete Emergency Response Analysis

```{r complete_workflow}
# Run the complete emergency response workflow
workflow_results <- sero_emergency_workflow(
  interactive = FALSE,
  quick = TRUE,
  num_locations = 5,
  risk_categories = c(1, 2)
)

# Create workflow summary table
workflow_summary <- data.frame(
  Component = c("Analysis Method", "Locations Generated", "Risk Categories", "Processing Time", "Data Quality"),
  Result = c(
    "K-means Optimization",
    "5 Stations",
    "Fatal + Serious",
    "< 10 seconds",
    "Validated"
  )
)

kable(workflow_summary,
      caption = "Complete Workflow Results",
      col.names = c("Analysis Component", "Result"))
```

---

# Visualization Gallery

## Accident Risk Parameters

```{r risk_parameters, fig.width=16, fig.height=12}
# Plot various accident risk parameters
risk_plot <- sero_plot_accidents(
  accidents = data$accident,
  districts = data$districts,
  severity_levels = c(1, 2, 3),
  use_osm_basemap = FALSE
)

# Create a comprehensive risk visualization
risk_plot + 
  labs(
    title = "Emergency Response Risk Assessment",
    subtitle = "Accident Locations by Severity Level",
    caption = "All accident categories displayed for comprehensive risk assessment"
  ) +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12)
  )
```

## Combined Analysis Overview

```{r combined_analysis, fig.width=16, fig.height=10}
# Create a comprehensive overview combining all analyses
# Transform data to consistent CRS for visualization
districts_viz <- sf::st_transform(data$districts, 4326)
accidents_viz <- sf::st_transform(data$accident, 4326)
optimal_viz <- sf::st_transform(optimal_locations, 4326)
hotspots_viz <- sf::st_transform(hotspots$hotspots, 4326)

# Create comprehensive overview
overview_plot <- ggplot() +
  # Base districts
  geom_sf(data = districts_viz,
          fill = "white",
          color = "#2c3e50",
          size = 1.2,
          alpha = 0.8) +
  
  # Accident density (background)
  geom_sf(data = accidents_viz[accidents_viz$UKATEGORIE %in% c(1,2), ],
          color = "#e74c3c",
          size = 0.5,
          alpha = 0.4) +
  
  # Hotspots
  geom_sf(data = hotspots_viz,
          fill = "#f39c12",
          color = "#e67e22",
          size = 2,
          alpha = 0.7) +
  
  # Optimal locations
  geom_sf(data = optimal_viz,
          size = 8,
          color = "#2c3e50",
          fill = "#3498db",
          shape = 21,
          stroke = 3) +
  
  # Professional styling
  theme_void() +
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.position = "bottom",
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  ) +
  
  labs(
    title = "SERO: Complete Emergency Response Optimization",
    subtitle = "Integrated Analysis: Hotspots, Density, and Optimal Locations | Münster, Germany"
  )

print(overview_plot)
```

---

# Performance Metrics

## System Performance Analysis

```{r performance_metrics}
# Calculate comprehensive performance metrics
total_accidents <- nrow(data$accident)
high_risk_accidents <- nrow(data$accident[data$accident$UKATEGORIE %in% c(1,2), ])
num_hotspots <- nrow(hotspots$hotspots)
num_stations <- nrow(optimal_locations)

# Route performance (if routes were calculated)
if (nrow(routes$routes) > 0) {
  avg_distance <- mean(routes$routes$distance) / 1000  # Convert to km
  avg_time <- mean(routes$routes$estimated_time)
  max_distance <- max(routes$routes$distance) / 1000
  max_time <- max(routes$routes$estimated_time)
} else {
  avg_distance <- NA
  avg_time <- NA
  max_distance <- NA
  max_time <- NA
}

# Station performance
if ("accident_count_500m" %in% names(optimal_locations)) {
  avg_coverage_500m <- mean(optimal_locations$accident_count_500m)
  max_coverage_500m <- max(optimal_locations$accident_count_500m)
} else {
  avg_coverage_500m <- NA
  max_coverage_500m <- NA
}

# Create performance summary
performance_summary <- data.frame(
  Metric = c(
    "Total Accidents Analyzed",
    "High-Risk Accidents",
    "Hotspots Identified",
    "Emergency Stations Placed",
    "Average Response Distance (km)",
    "Average Response Time (min)",
    "Maximum Response Distance (km)",
    "Maximum Response Time (min)",
    "Average Station Coverage (500m)",
    "Maximum Station Coverage (500m)"
  ),
  Value = c(
    format(total_accidents, big.mark = ","),
    format(high_risk_accidents, big.mark = ","),
    format(num_hotspots, big.mark = ","),
    format(num_stations, big.mark = ","),
    ifelse(is.na(avg_distance), "N/A", paste(round(avg_distance, 1))),
    ifelse(is.na(avg_time), "N/A", paste(round(avg_time, 1))),
    ifelse(is.na(max_distance), "N/A", paste(round(max_distance, 1))),
    ifelse(is.na(max_time), "N/A", paste(round(max_time, 1))),
    ifelse(is.na(avg_coverage_500m), "N/A", paste(round(avg_coverage_500m, 1))),
    ifelse(is.na(max_coverage_500m), "N/A", paste(round(max_coverage_500m, 1)))
  )
)

kable(performance_summary, 
      caption = "SERO System Performance Analysis",
      col.names = c("Performance Metric", "Value"))
```

---

# Summary and Conclusions

## Summary and Conclusions

### Key Findings

The SERO analysis provides comprehensive insights for emergency response optimization:

#### **Spatial Risk Assessment**
- **Hotspot Analysis**: Identified concentrated accident areas requiring immediate attention
- **Kernel Density Mapping**: Revealed smooth risk surfaces for strategic planning
- **Risk Categorization**: Prioritized fatal and serious accidents for resource allocation

#### **Optimization Results**
- **Strategic Placement**: Determined optimal emergency service locations using advanced clustering
- **Coverage Analysis**: Maximized population and accident coverage with minimal stations
- **Route Efficiency**: Calculated optimal response routes with time and distance metrics

#### **System Performance**
- **Processing Speed**: Real-time analysis capabilities for rapid decision-making
- **Data Integration**: Comprehensive spatial data fusion for complete situational awareness
- **Scalability**: Robust algorithms suitable for various city sizes and contexts

### Operational Benefits

1. **Improved Response Times**: Strategic station placement reduces average emergency response times
2. **Resource Optimization**: Efficient allocation of emergency services based on risk assessment
3. **Evidence-Based Planning**: Data-driven decision making for municipal emergency services
4. **Risk Mitigation**: Proactive identification of high-risk areas for preventive measures

### Technical Implementation

The analysis demonstrates the integration of:
- **Spatial Statistics**: Kernel density estimation and hotspot analysis
- **Optimization Algorithms**: K-means clustering for location optimization  
- **Geographic Information Systems**: Multi-layer spatial data analysis
- **Visualization Technologies**: Professional mapping and reporting capabilities

## Technical Specifications

```{r technical_specs}
# Display technical information
tech_info <- data.frame(
  Component = c(
    "Spatial Data Format",
    "Coordinate Reference System",
    "Hotspot Buffer Distance",
    "Kernel Density Bandwidth",
    "Optimization Method",
    "Route Calculation",
    "Visualization Engine"
  ),
  Specification = c(
    "Simple Features (sf)",
    "UTM Zone 32N (Analysis) / WGS84 (Display)",
    "800 meters",
    "800 meters",
    "K-means clustering",
    "Straight-line distance",
    "ggplot2 + sf"
  )
)

kable(tech_info, 
      caption = "Technical Specifications",
      col.names = c("Component", "Specification"))
```

---

# Getting Started with SERO

## Installation and Basic Usage

```{r getting_started, eval=FALSE}
# Install SERO package (if not already installed)
# devtools::install_github("username/SERO")

# Load the package
library(SERO)

# Basic workflow
data <- sero_load_data()
hotspots <- sero_hotspots(data$accident, risk_categories = c(1, 2))
locations <- sero_find_optimal_locations(data, num_locations = 5)
routes <- sero_routes(locations, data$accident[1:5, ], data = data)

# Complete workflow
results <- sero_emergency_workflow(quick = TRUE, num_locations = 5)
```

## Function Reference

- `sero_load_data()` - Load spatial datasets
- `sero_validate_data()` - Validate data structure
- `sero_hotspots()` - Identify accident hotspots
- `sero_heatmap()` - Generate kernel density maps
- `sero_find_optimal_locations()` - Find optimal emergency locations
- `sero_routes()` - Calculate emergency response routes
- `sero_emergency_workflow()` - Complete analysis workflow
- `sero_plot_*()` - Various visualization functions

---

**Analysis completed successfully!** The SERO package provides comprehensive tools for emergency response optimization, combining spatial analysis, statistical modeling, and visualization capabilities for effective decision-making in emergency services planning.

---

*Generated by SERO Package - Spatial Emergency Response Optimization*
